---
title: "africountrynames-testing"
output: html_document
---

Experimenting with African countrynames, planning to put into own micro package (e.g. to avoid repetition between other afrimapr packages, afriadmin and afrihealthsites).

ISSUE :
Always seems to be minor differences in some African countrynames between different datasets. To be absolutely sure which names are used create a package containing a definitive list of names and their equivalent iso3c codes.

e.g. even the countrycode package returns some names from iso3c that aren't expected.

africountrynames package can also serve as example of as minimal package that can be used in teaching for the afrimapr community.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode)

#remotes::install_github("afrimapr/afriadmin")
library(afriadmin) # partly just to get at country list, also potential to compare with other boundaries
library(mapview)
library(knitr) #for kable
library(dplyr)
library(ggplot2)

# install.packages("remotes")
remotes::install_gitlab("dickoa/rgeoboundaries")
library(rgeoboundaries)

```


```{r, eval=TRUE, warning=FALSE}


#demo that taking the list of iso3c codes from afriadmin and converting them with package countrycode to names
#results in some that aren't what you might want

#todo rename afcountries
data(afcountries)

#countrycode(afcountries$iso3c, origin = 'iso3c', destination = 'iso.name')

countrycode(afcountries$iso3c, origin = 'iso3c', destination = 'country.name')

#countrycode country.name has "Congo - Kinshasa"         "Congo - Brazzaville"

#useful for me
#countrycode('rwanda','country.name','iso3c') #[1] "RWA"
#idea should suggest or pull request that countrycode adds a function iso3c() that just has countryname arg.
#otherwise need to remember country.name

# can use the codelist from package countrycode, which contains 59 African countries
# and 743 columns of different codes, names, languages, regex etc.

dfaf <- countrycode::codelist[which(codelist$continent=="Africa"),]

#country.name.en has the weird Congos
#cow.name is the closest to what I have (only maybe problem is that it doesn't have Western Sahara (ESH) - but that doesn't appear in many things anyway)
# has no accents, Ivory Coast, Congo, Tanzania

countrycode(afcountries$iso3c, origin = 'iso3c', destination = 'cow.name')

# I think we don't need africountrynames package - unless maybe we use it as a wrapper to countrycode

# for other afrimapr packages, i think we should work with iso3c, convert everything to that & then convert to cow.name when we need

# also maybe subset out those that don't have cow.name 
dfaf_noiso <- countrycode::codelist[which(codelist$continent=="Africa" & is.na(codelist$cow.name)),]

dfaf_noiso$country.name.en
#[1] "Mayotte"        "Réunion"        "Somaliland"     "St. Helena"     "Western Sahara"

# can potentially sort western sahara using countrycode::custom_match
custom_match <- c(ESH = "Western Sahara")

countrycode("ESH", origin = 'iso3c', destination = 'cow.name', custom_match = custom_match)



```




