---
title: "africountrynames-testing"
output: html_document
---

Experimenting with African countrynames, planning to put into own micro package (e.g. to avoid repetition between other afrimapr packages, afriadmin and afrihealthsites).

ISSUE :
Always seems to be minor differences in some African countrynames between different datasets. To be absolutely sure which names are used create a package containing a definitive list of names and their equivalent iso3c codes.

e.g. even the countrycode package returns some names from iso3c that aren't expected.

africountrynames package can also serve as example of as minimal package that can be used in teaching for the afrimapr community.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode)

library(afriadmin) # partly just to get at country list, also potential to compare with other boundaries
library(mapview)
library(knitr) #for kable
library(dplyr)
library(ggplot2)

# install.packages("remotes")
remotes::install_gitlab("dickoa/rgeoboundaries")
library(rgeoboundaries)

```


```{r, eval=TRUE, warning=FALSE}


#demo that taking the list of iso3c codes from afriadmin and converting them with package countrycode to names
#results in some that aren't what you might want

#todo rename afcountries
data(afcountries)

#countrycode(afcountries$iso3c, origin = 'iso3c', destination = 'iso.name')

countrycode(afcountries$iso3c, origin = 'iso3c', destination = 'country.name')

#countrycode country.name has "Congo - Kinshasa"         "Congo - Brazzaville"



```




