---
title: "compare_healthsites_kemri"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(afrihealthsites)
library(knitr) #for kable
library(dplyr)
library(ggplot2)

```


```{r, eval=FALSE, echo=FALSE, warning=FALSE}

#eval = FALSE because run already and takes few mins

  dfallcountries <- NULL

  #takes a few minutes to do all countries
  for( country in afcountries$name)
  {
    
    dfsumm <- merge_points(country, toreturn='summary', 
                           hs_amenity=c('clinic', 'doctors', 'pharmacy', 'hospital'),
                           dist_same_m = 50)
    
    dfallcountries <- rbind(dfallcountries, dfsumm)
  }

#reformatting dataframe

# rename columns containing num locations from first entry in source columns
names(dfallcountries)[which(names(dfallcountries)=='numpoints1')] <- as.character(dfallcountries$source1[1])
names(dfallcountries)[which(names(dfallcountries)=='numpoints2')] <- as.character(dfallcountries$source2[1])

# copy and rename object
df_hs_who_compare_50m <- dfallcountries

# remove source columns
df_hs_who_compare_50m <- df_hs_who_compare_50m[,-c(2,4)]

# save object
save(df_hs_who_compare_50m, file="data//df_hs_who_compare_50m.rda") 

```

```{r table_compare, echo=FALSE, asis=TRUE}

load("data//df_hs_who_compare_50m.rda")

#TODO check seeming error with merging

#miss out threshdist column (4)
knitr::kable(df_hs_who_compare_50m[,-4], caption="comparing healthsites and KEMRI collated data")

#can I output the table as a png or similar


```


```{r bar_compare, eval=FALSE,echo=FALSE, asis=TRUE}

df2 <- dplyr::arrange(df_hs_who_compare_50m, num_merged)

positions <- df2$country

ggplot(df2, aes(x=country, y=num_merged, label=num_merged)) + 
  geom_bar(stat='identity', width=.5)  + #, aes(fill=mpg_type)
  # scale_fill_manual(name="Mileage", 
  #                   labels = c("Above Average", "Below Average"), 
  #                   values = c("above"="#00ba38", "below"="#f8766d")) + 
  # labs(subtitle="Normalised mileage from 'mtcars'", 
  #      title= "Diverging Bars") +
  scale_x_discrete(limits = positions) +
  coord_flip()

```

Numbers of locations from each data source and locations that are shared based on proximity. 

```{r points_numbers_compare, echo=FALSE, warnings=FALSE, asis=TRUE, fig.width=9, fig.height=9}

# arrange in order of number merged
df2 <- dplyr::arrange(df_hs_who_compare_50m, num_merged)

positions <- df2$country

# ggplot(df2, aes(x=country, y=num_merged, label=num_merged)) +
#   geom_point(fill=NA)  + #, aes(fill=mpg_type)
#   geom_point(aes(y=who), col='steelblue2')  +
#   geom_point(aes(y=healthsites), col='firebrick1')  +
#   geom_point(aes(y=num_shared), col='darkgreen', shape='+', size=2)  +
#   ylab("health facility locations from different sources") +
#   scale_x_discrete(limits = positions) +
#   theme_minimal() +
#   coord_flip()

# alternative tidyer method using pivot_longer() allows legend

# pivot longer to structure data for plot
df3 <- tidyr::pivot_longer(df2, -c(country, threshdistm), names_to = "measure", values_to = "count")

# filter just rows wanted in plot
df4 <- dplyr::filter(df3,measure %in% c('healthsites','who','num_shared','num_merged'))

# set factor order otherwise colours are wrong
df4$measure <- factor(df4$measure,levels=c("who", "healthsites", "num_shared", "num_merged"))

# set colours
point_cols <- c("who"='steelblue2', "healthsites"='firebrick1', "num_shared"='green', "num_merged"='black')

ggplot(df4, aes(x=country, y=count, colour=measure)) + 
  geom_point(alpha = 0.5) +
  scale_colour_manual(name="data source",
                    labels = c("WHO-KEMRI", "healthsites.io","shared","merged"),
                    values = point_cols) +
  # labs(subtitle="Normalised mileage from 'mtcars'", 
  #      title= "Diverging Bars") +
  ylab("number of health facility locations") +
  scale_x_discrete(limits = positions) +
  #scale_y_log10() +
  theme_minimal() +
  coord_flip()

```

Numbers of locations from each data source and locations that are shared based on proximity. 
Log scale to show for smaller countries too.

```{r points_numbers_log_compare, echo=FALSE, warnings=FALSE, asis=TRUE, fig.width=9, fig.height=9}

# log version
ggplot(df4, aes(x=country, y=count, colour=measure)) + 
  geom_point(alpha = 0.5) +
  scale_colour_manual(name="data source",
                    labels = c("WHO-KEMRI", "healthsites.io","shared","merged"),
                    values = point_cols) +
  # labs(subtitle="Normalised mileage from 'mtcars'", 
  #      title= "Diverging Bars") +
  ylab("number of health facility locations") +
  scale_x_discrete(limits = positions) +
  scale_y_log10(limits=c(50,30000)) +
  #ylim(c(10,15000)) +
  theme_minimal() +
  coord_flip()
  
```

Percentage contribution of locations from each source to a merged dataset.  

```{r percentages_compare, echo=FALSE, warnings=FALSE, asis=TRUE, fig.width=9, fig.height=9}

# plot a point for each source and it's % of the maximum 
# actually easier and maybe clearer way of doing is percentage of num-merged 

df2 <- df_hs_who_compare_50m

df2$perc_who_merg <- 100 * df2$who/df2$num_merged
df2$perc_hs_merg <- 100 * df2$healthsites/df2$num_merged

# arrange in order of % in who
df2 <- dplyr::arrange(df2, perc_who_merg)

positions <- df2$country

# ggplot(df2, aes(x=country, y=num_merged, label=num_merged)) + 
#   #geom_point(fill=NA)  + #, aes(fill=mpg_type)
#   geom_point(aes(y=perc_who_merg), col='steelblue2')  + 
#   geom_point(aes(y=perc_hs_merg), col='firebrick1')  +
#   ylab("percentage of locations from healthsites(red) & WHO-Kemri(blue) making up a merged dataset") +
#   scale_x_discrete(limits = positions) +
#   theme_minimal() +
#   coord_flip()

# alternative tidyer method using pivot_longer() allows legend

df3 <- tidyr::pivot_longer(df2, -c(country, threshdistm), names_to = "measure", values_to = "count")

df4 <- dplyr::filter(df3,measure %in% c('perc_who_merg','perc_hs_merg'))

ggplot(df4, aes(x=country, y=count, colour=measure)) + 
  geom_point() +
  scale_colour_manual(name="data source",
                    labels = c("WHO-KEMRI", "healthsites.io"),
                    values = c("perc_who_merg"='steelblue2', "perc_hs_merg"='firebrick1')) +
  # labs(subtitle="Normalised mileage from 'mtcars'", 
  #      title= "Diverging Bars") +
  ylab("percentage of locations making up a merged dataset") +
  scale_x_discrete(limits = positions) +
  theme_minimal() +
  coord_flip()

```

Compare per country distributions of facility types between who and healthsites 

```{r compare_type_frequencies, echo=FALSE, warnings=FALSE, asis=TRUE, fig.width=9, fig.height=9}

# plan to add this as a function to healthsites

# start with function to work on a single source

country <- 'Togo'
type_filter = c('clinic', 'doctors', 'pharmacy', 'hospital')
type_filter = 'all'
datasource = 'healthsites' #'who'

sf1 <- afrihealthsites(country, datasource = datasource, plot=FALSE, hs_amenity=type_filter) #, who_type=who_type)

tot_facilities <- nrow(sf1)

# ggplot(sf1) + 
#   geom_bar(aes(y = amenity)) 

# add text labels of frequency
ggplot(sf1, aes(y = amenity, fill = amenity)) + 
  geom_bar(show.legend=FALSE) + 
  theme_minimal() +
  geom_text(stat='count', aes(label=..count..), hjust=-1) +
  labs(title=paste("Selected facility types from", datasource), 
       subtitle= paste("Total =",tot_facilities)) +
  NULL

#TODO repeat for all countries
#add to function
#add to shiny app
#sum the numbers for all categories & display somehow


```


