---
title: "Comparing Kenyan Health facility location data from different sources"
#output: html_document
author: "Andy South @afrimapr"
date: "`r Sys.Date()`"
output: pdf_document
---

DRAFT
Code here:[https://github.com/afrimapr/afrimapr_dev/blob/master/compare_moh_kenya.Rmd](https://github.com/afrimapr/afrimapr_dev/blob/master/compare_moh_kenya.Rmd)

We currently have access to the following sources of health facility locations for Kenya.

1. downloaded from Ministry of Health website (no coordinates) [http://kmhfl.health.go.ke/#/facility_filter/results](http://kmhfl.health.go.ke/#/facility_filter/results)
1. a file created by WiGISKe 2021-04 from multiple sources
1. healthsites.io in OpenStreetMap
1. collated data from published paper Maina(2019) 

Following discussion 2021-04-29 suggestion that modelling work at the ward level uses the MoH data that has advantage of having a clear provenance (and also has the most facilities and attributes).

Also note that the MoH data has a column named 'Code' which matches with the 'MFL CODE' column in the pdf of MoH approved Vaccination posts so will allow us to link the two. 

Numbers of locations per source.

source  | number facilities | spatial data | useful attributes
-------- | -------------- | ------------------ | ------------------
1. MoH     |  12378 | Ward | Registration_number,Keph level,Owner type,Beds,Cots,Operation status,Open_weekends
2. WiGISKe |  10013  | coords | Owner
3. OSM    | 2032 | coords | (few complete) beds,staff_doctors,staff nurses


```{r setup, include=FALSE}

library(afrihealthsites)
library(knitr) #for kable
library(dplyr)
library(ggplot2)
library(sf) #to read in shapefile


knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(global.par = TRUE)

```



```{r, include=FALSE}
#to stop fig margins getting cut
#https://stackoverflow.com/questions/32827651/plot-margins-in-rmarkdown-knitr
par(mar=c(5,5,5,5)) #it's important to have that in a separate chunk
```


```{r read-wigiske, eval=TRUE, include=FALSE}

file_wigiske <- "data-raw\\health-facilities-kenya-2021-04\\Healthcare_Facilities.shp"

# test reading in file
sfwigiske <- sf::st_read(file_wigiske)

names(sfwigiske)
# [1] "OBJECTID"   "Facility_N" "Type"       "Owner"      "County"     "Sub_County" "Division"   "Location"  
# [9] "Sub_Locati" "Constituen" "Nearest_To" "Latitude"   "Longitude"  "geometry"  

length(names(sfwigiske))
#14

#26 types
types_wigiske <- unique(sfwigiske[['Type']])
#22 owners
owners_wigiske <- unique(sfwigiske[['Owner']])

nrow(sfwigiske)
# 10013



```



```{r read-moh, eval=TRUE, include=FALSE}

file_moh <- "data-raw\\health-facilities-moh-kenya.csv"

# test reading in file
dfmoh <- read.csv(file_moh, as.is=TRUE, check.names=FALSE)

names(dfmoh)
#  [1] "Code"                   "Name"                   "Officialname"           "Registration_number"   
#  [5] "Keph level"             "Facility type"          "Facility_type_category" "Owner"                 
#  [9] "Owner type"             "Regulatory body"        "Beds"                   "Cots"                  
# [13] "County"                 "Constituency"           "Sub county"             "Ward"                  
# [17] "Operation status"       "Open_whole_day"         "Open_public_holidays"   "Open_weekends"         
# [21] "Open_late_night"        "Service_names"          "Approved"               "Public visible"        
# [25] "Closed" 

#length(names(dfmoh))
#25

#29 cats
fac_types_moh <- unique(dfmoh[['Facility type']])
#9 cats
fac_type_cats_moh <- unique(dfmoh[['Facility_type_category']])
#19 owners
owners_moh <- unique(dfmoh[['Owner']])

nrow(dfmoh)
# 12378

```

```{r read-osm, eval=TRUE, include=FALSE}


# cached in afrihealthsites 1593 facilities
dfosm <- afrihealthsites('kenya',datasource = 'healthsites')
# osm live 2032 facilities
#dfosm <- afrihealthsites('kenya',datasource = 'healthsites_live')

#names(dfosm)
#  [1] "osm_id"              "osm_type"            "completeness"        "is_in_health_zone"   "amenity"          
#  [6] "speciality"          "addr_street"         "part_time_beds"      "tag"                 "operator"         
# [11] "water_source"        "changeset_id"        "insurance"           "staff_doctors"       "contact_number"   
# [16] "uuid"                "electricity"         "opening_hours"       "operational_status"  "source"           
# [21] "is_in_health_area"   "health_amenity_type" "changeset_version"   "emergency"           "changeset_timestamp"
# [26] "addr_housenumber"    "addr_postcode"       "addr_city"           "name"                "staff_nurses"     
# [31] "changeset_user"      "wheelchair"          "beds"                "url"                 "dispensing"       
# [36] "healthcare"          "operator_type"       "geometry" 

length(names(dfosm))
#38

#6 types
types_osm <- unique(dfosm[['amenity']])
#or 
types2_osm <- unique(dfosm[['healthcare']])
#21 owners
owners_moh <- unique(dfosm[['operator_type']])

nrow(dfosm)
# 2032

```

## Comparing frequency of facility types in MoH & WiGISKe data

```{r compare_types_moh_wigiske, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=8, fig.height=9}


    library(patchwork)

    gg1 <- afrihealthsites::facility_types("kenya", 
                                           datasource=file_moh,
                                           type_column = 'Facility type',
                                           lonlat_columns = NULL,
                                           plot=FALSE,
                                           plot_title = "MoH data : Frequency of facility types (no coords)")

    gg2 <- afrihealthsites::facility_types("kenya", 
                                           datasource=sfwigiske,
                                           type_column = 'Type',
                                           plot=FALSE,
                                           plot_title = "WiGISKe data : Frequency of facility types (with coords)")

    

      #set xmax to be the same for both plots
      #hack to find max xlim for each object
      #TODO make this less hacky ! it will probably fail when ggplot changes
      max_x1 <- max(ggplot_build(gg1)$layout$panel_params[[1]]$x$continuous_range)
      max_x2 <- max(ggplot_build(gg2)$layout$panel_params[[1]]$x$continuous_range)
      #set xmax for both plots to this
      gg1 <- gg1 + xlim(c(0,max(max_x1,max_x2, na.rm=TRUE)))
      gg2 <- gg2 + xlim(c(0,max(max_x1,max_x2, na.rm=TRUE)))

      #set size of y plots to be dependent on num cats
      #y axis has cats, this actually gets max of y axis, e.g. for 6 cats is 6.6
      max_y1 <- max(ggplot_build(gg1)$layout$panel_params[[1]]$y$continuous_range)
      max_y2 <- max(ggplot_build(gg2)$layout$panel_params[[1]]$y$continuous_range)

      #setting heights to num cats makes bar widths constant between cats
      gg1 / gg2 + plot_layout(heights=c(max_y1, max_y2)) #patchwork



```

## MoH facility attributes

```{r moh_facility_type_category, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=5}

# Facility_type_category
ggken <- facility_types("kenya", 
                        datasource=dfmoh,
                        type_column = 'Facility_type_category',
                        lonlat_columns = NULL,
                        #plot=FALSE,
                        plot_title = "MoH data : Frequency of facility type categories (no coords)")

#ggken  + plot_layout(ncol=2,widths=c(1,0.1))
```


```{r moh_keph_level, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=4}

# Keph level, defined facility levels
ggken <- facility_types("kenya", 
                        datasource=dfmoh,
                        type_column = "Keph level",
                        lonlat_columns = NULL,
                        plot_title = "MoH data : Keph level (no coords)")
#plot(ggken)
```


```{r moh_owner_type, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=3}

# Owner type = public/private
ggken <- facility_types("kenya", 
                        datasource=dfmoh,
                        type_column = "Owner type",
                        lonlat_columns = NULL,                        
                        plot_title = "MoH data : Owner type (no coords)")
#plot(ggken)


#as.data.frame(table(dfmoh$`Owner type`))
# 1       Faith Based Organization 1035
# 2             Ministry of Health 5875
# 3 Non-Governmental Organizations  366
# 4               Private Practice 5102


#number of beds - but the default plot doesn't cope well with the numeric data
# ggken <- facility_types("kenya", 
#                         datasource=dfmoh,
#                         type_column = 'Beds',
#                         lonlat_columns = NULL,                        
#                         plot_title = "MoH data : Number of beds (no coords)")
# plot(ggken)



```



```{r owner-wigiske, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=6}


#22 owners
owners_wigiske <- unique(sfwigiske[['Owner']])

ggken <- facility_types("kenya", 
                        datasource=sfwigiske,
                        type_column = "Owner",
                        plot_title = "WiGISKe data : Owner")


```

## All attribute columns in WiGISKe & MoH data 

```{r attribute columns, echo=FALSE, warning=FALSE}

#WiGISKe
names(sfwigiske)

#MoH
names(dfmoh)

```


```{r map_moh_kenya, eval=FALSE, include=FALSE}

#eval=FALSE because no coordinates

# just view MoH data
sfmoh <- afrihealthsites("kenya", datasource=file_moh,
                             type_column = 'Facility_type_category',
                             label_column = 'Facility Name',
                             lonlat_columns = c("Longitude", "Latitude"))

#testing filter
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter='Dispensary')
#filter with two types
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter=c('Dispensary','Health Center'))

#trying to remove one from selection doesn't work
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter='-Dispensary')
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter='!Dispensary')

# plot moh vs kemri on a map
# this does work now but only because the defaults are set up for kenya
# todo sort code so will work for other countries
compare_hs_sources("kenya", datasources=c('who',file_moh),
                   type_column = 'Facility Type',
                   label_column = 'Facility Name')
                   #lonlat_columns = c("Longitude", "Latitude"))

```

# Number of facilities per ward in MoH data
```{r moh-per-ward, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=6}

library(dplyr)

dfcountbyward <- dfmoh %>% group_by(Ward) %>% summarise(count=n())

ggplot(dfcountbyward, aes(x=count)) + geom_bar()

```


```{r moh_type_and_owner_type, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=4}

# bar chart with columns for 'Facility_type_category', coloured by "Owner type"

#note 2 spaces after health
toexclude <- c("Primary health  care services", "None")
dfmoh2 <- dfmoh[-which(dfmoh$Facility_type_category %in% toexclude ),]

gg <- ggplot2::ggplot(dfmoh2, aes(y = Facility_type_category, fill = `Owner type`))

gg <- gg + geom_bar(show.legend=TRUE) +
      labs(title="MoH data - ownership by facility type") +
      theme_minimal() 
      
gg

```

```{r moh_kephlevel_and_owner_type, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=4}


gg <- ggplot2::ggplot(dfmoh, aes(y = `Keph level`, fill = `Owner type`))

gg <- gg + geom_bar(show.legend=TRUE) +
      labs(title="MoH data - ownership by Keph level") +
      theme_minimal() 
      
gg

```

Draft code to :

1. scrape vaccination centre locations from PDF 
1. join to MoH Master facility list to get ward names
1. join to wards map

```{r moh_vaccination_centres, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE, asis=TRUE, fig.width=11, fig.height=10}

# looking at getting vaccination sites from pdf
# https://www.brodrigues.co/blog/2018-06-10-scraping_pdfs/

library(pdftools)
library(stringr) #string manipulation
library(mapview) #interactive mapping
library(tmap) #ststic maps

download.file(url = "https://www.health.go.ke/wp-content/uploads/2021/03/MOH-APPROVED-COVID-19-VACCINATION-POSTS-MARCH-TO-JUNE_FINAL.pdf",
                   destfile = "data-raw/ken-vacc-posts.pdf", mode = "wb")

filename = "data-raw/ken-vacc-posts.pdf"

#puts the text into 9 char strings, 1 per page
txt <- pdf_text(filename)

#this splits page 1 into lines, puts each into a separate column
txt1 <- stringr::str_split(txt[1], "\n", simplify = TRUE)

#str(txt1)
#chr [1, 1:78]

#on page1 line5 contains the headers
#txt1[1,5]
#[1] "                        Subcounty                    Health Facility Name             MFL CODE    Ownership\r"
#on page1 lines6 & after contain the data
#luckily columns are separated by multiple spaces
#txt1[1,6]
#[1] "1               Mvita                  Coast General Teaching and Referal Hospital    11289    Public\r"
#beware some lines contain extra word that is the county
#should be relatively easy to remove later
#txt1[1,10]
#[1] "5   Mombasa Mvita                      Aga Khan Hospital Mombasa                      11203    Private\r"

#removing all countynames followed by 2 spaces e.g. 'Kwale '
#to avoid problem caused by column2 County names which contains entries for only a few rows
countyregex <- c("Kwale  |Kilifi  |Tana River  |Lamu  |Taita Taveta  |Garissa  |Wajir  |Mandera  " )
txt1 <- str_remove_all(txt1, countyregex)
#txt1 <- str_remove_all(txt1, c("Kwale  "))
#replace more than 2 spaces with '|' then use that as a delimiter to read into a dataframe
txt1columns <- str_replace_all(txt1, "\\s{2,}", "|")


#remove first 4 rows from page1
txt1columns2 <- txt1columns[-c(1:4)]
#add extra item to header row
txt1columns2[1] <- paste("vacc_center_num",txt1columns2[1])
#make text connection to the object
text_con <- textConnection(txt1columns2)
#read into a dataframe
df1 <- read.csv(text_con, sep = "|") #, skip=4)
#df1 <- readr::read_delim(text_con, delim = "|", skip=4)

#mostly works - this issue now mostly fixed above
#issue with this line where the county (Kwale) gets read in as an extra column
#problem because the county column has cells merged across multiple rows
#I could fix by removing counties at the txt or txt1 stage
#would cause county names to be removed from facility names e.g. Mombasa hospital
#but the MFL code would still allow joining
#could fix by removing e.g. '|Kwale', but that could cause problems by removing later columns
#removing all countynames followed by 2 spaces should sort it e.g. 'Kwale '
#another way could be to count number of | and if > 4 remove the first
#str_count(txt1columns2,"\\|")
#4 4 4 4 4 4 4 4 4 4 4 4 4 5 4 4 4 4 4 4 4 4 4 5 4 4 4 4 4 4 4 4 4 4 5 4 4 4 4 1 4 4 4 4 4 4 4 4 4 1 4 4 4 3 4 4 4 4 4 1 4 4 4 4 4 4 4 4 4 4 2 2 1 0
#txt1columns2[18]
#[1] "13|Kwale|Lungalunga|Lungalunga Sub County Hospital|11526|Public\r"


# checking whether joining this onto the MoH data works
# try left join df1,MoH so it is easier to see
# this does work for ~70 facilities, a few problems
dfjoined <- dplyr::left_join(x=df1, y=dfmoh,
                             by=c("MFL.CODE"="Code")
                            )

#try to join to a ward map to test

#read wards map (downloaded from HDX), could get via rhdx
sfwards <- sf::st_read("data-raw/kenya-wards-hdx/kenya_wards.shp")

#remove " Ward" from the ward names in sfwards$ward to allow joining
sfwards$ward2 <- stringr::str_remove(sfwards$ward," Ward")

#join vaccination posts to wards map
#Some fail to join probably due to differences in ward naming
sfwards_vacc <- dplyr::left_join(x=dfjoined, y=sfwards,
                             by=c("Ward"="ward2")
                            )
#re sf the object, which seems to forget it is sf in the join
sfwards_vacc <- sf::st_as_sf(sfwards_vacc)

#view the wards that contain a facility on page1 
#mapview::mapview(sfwards_vacc, label="Health.Facility.Name")

#static map
tmap::tm_shape(sfwards_vacc) +
    tm_borders("blue", lwd = .5) +
    tm_layout(main.title="Wards containing vaccination posts on page 1 of MoH pdf",
              main.title.size=1) +
    tm_text("Health.Facility.Name", size=0.5) 


```

#TODO 
#work out how to deal with few remaining rows that fail on page1
#check on wards that fail to join
#try applying to later pages 
