---
title: "Comparing Kenyan Health facility location data from different sources"
#output: html_document
author: "Andy South @afrimapr"
date: "`r Sys.Date()`"
output: pdf_document
---

DRAFT
Code here:[https://github.com/afrimapr/afrimapr_dev/blob/master/compare_moh_kenya.Rmd](https://github.com/afrimapr/afrimapr_dev/blob/master/compare_moh_kenya.Rmd)

We currently have access to the following sources of health facility locations for Kenya.

1. downloaded from Ministry of Health website (no coordinates) [http://kmhfl.health.go.ke/#/facility_filter/results](http://kmhfl.health.go.ke/#/facility_filter/results)
1. a file created by WiGISKe 2021-04 from multiple sources
1. healthsites.io in OpenStreetMap
1. collated data from published paper Maina(2019) 

We will initially concentrate on comparing the first 3, given that we know the Maina(2019) data are older.

Following discussion 2021-04-29 suggestion that modelling work at the ward level uses the MoH data that has advantage of having a clear provenance (and also has the most facilities and attributes).

Numbers of locations per source.

source  | number facilities | spatial data | useful attributes
-------- | -------------- | ------------------ | ------------------
1. MoH     |  12378 | Ward | Registration_number,Keph level,Owner type,Beds,Cots,Operation status,Open_weekends
2. WiGISKe |  10013  | coords | Owner
3. OSM    | 2032 | coords | (few complete) beds,staff_doctors,staff nurses


```{r setup, include=FALSE}

library(afrihealthsites)
library(knitr) #for kable
library(dplyr)
library(ggplot2)
library(sf) #to read in shapefile

knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(global.par = TRUE)

```



```{r, include=FALSE}
#to stop fig margins getting cut
#https://stackoverflow.com/questions/32827651/plot-margins-in-rmarkdown-knitr
par(mar=c(5,5,5,5)) #it's important to have that in a separate chunk
```


```{r read-wigiske, eval=TRUE, include=FALSE}

file_wigiske <- "data-raw\\health-facilities-kenya-2021-04\\Healthcare_Facilities.shp"

# test reading in file
sfwigiske <- sf::st_read(file_wigiske)

names(sfwigiske)
# [1] "OBJECTID"   "Facility_N" "Type"       "Owner"      "County"     "Sub_County" "Division"   "Location"  
# [9] "Sub_Locati" "Constituen" "Nearest_To" "Latitude"   "Longitude"  "geometry"  

length(names(sfwigiske))
#14

#26 types
types_wigiske <- unique(sfwigiske[['Type']])
#22 owners
owners_wigiske <- unique(sfwigiske[['Owner']])

nrow(sfwigiske)
# 10013



```



```{r read-moh, eval=TRUE, include=FALSE}

file_moh <- "data-raw\\health-facilities-moh-kenya.csv"

# test reading in file
dfmoh <- read.csv(file_moh, as.is=TRUE, check.names=FALSE)

names(dfmoh)
#  [1] "Code"                   "Name"                   "Officialname"           "Registration_number"   
#  [5] "Keph level"             "Facility type"          "Facility_type_category" "Owner"                 
#  [9] "Owner type"             "Regulatory body"        "Beds"                   "Cots"                  
# [13] "County"                 "Constituency"           "Sub county"             "Ward"                  
# [17] "Operation status"       "Open_whole_day"         "Open_public_holidays"   "Open_weekends"         
# [21] "Open_late_night"        "Service_names"          "Approved"               "Public visible"        
# [25] "Closed" 

#length(names(dfmoh))
#25

#29 cats
fac_types_moh <- unique(dfmoh[['Facility type']])
#9 cats
fac_type_cats_moh <- unique(dfmoh[['Facility_type_category']])
#19 owners
owners_moh <- unique(dfmoh[['Owner']])

nrow(dfmoh)
# 12378

```

```{r read-osm, eval=TRUE, include=FALSE}


# cached in afrihealthsites 1593 facilities
dfosm <- afrihealthsites('kenya',datasource = 'healthsites')
# osm live 2032 facilities
#dfosm <- afrihealthsites('kenya',datasource = 'healthsites_live')

#names(dfosm)
#  [1] "osm_id"              "osm_type"            "completeness"        "is_in_health_zone"   "amenity"          
#  [6] "speciality"          "addr_street"         "part_time_beds"      "tag"                 "operator"         
# [11] "water_source"        "changeset_id"        "insurance"           "staff_doctors"       "contact_number"   
# [16] "uuid"                "electricity"         "opening_hours"       "operational_status"  "source"           
# [21] "is_in_health_area"   "health_amenity_type" "changeset_version"   "emergency"           "changeset_timestamp"
# [26] "addr_housenumber"    "addr_postcode"       "addr_city"           "name"                "staff_nurses"     
# [31] "changeset_user"      "wheelchair"          "beds"                "url"                 "dispensing"       
# [36] "healthcare"          "operator_type"       "geometry" 

length(names(dfosm))
#38

#6 types
types_osm <- unique(dfosm[['amenity']])
#or 
types2_osm <- unique(dfosm[['healthcare']])
#21 owners
owners_moh <- unique(dfosm[['operator_type']])

nrow(dfosm)
# 2032

```

## Comparing frequency of facility types in MoH & WiGISKe data

```{r compare_types_moh_wigiske, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=8, fig.height=9}


    library(patchwork)

    gg1 <- afrihealthsites::facility_types("kenya", 
                                           datasource=file_moh,
                                           type_column = 'Facility type',
                                           lonlat_columns = NULL,
                                           plot=FALSE,
                                           plot_title = "MoH data : Frequency of facility types (no coords)")

    gg2 <- afrihealthsites::facility_types("kenya", 
                                           datasource=sfwigiske,
                                           type_column = 'Type',
                                           plot=FALSE,
                                           plot_title = "WiGISKe data : Frequency of facility types (with coords)")

    

      #set xmax to be the same for both plots
      #hack to find max xlim for each object
      #TODO make this less hacky ! it will probably fail when ggplot changes
      max_x1 <- max(ggplot_build(gg1)$layout$panel_params[[1]]$x$continuous_range)
      max_x2 <- max(ggplot_build(gg2)$layout$panel_params[[1]]$x$continuous_range)
      #set xmax for both plots to this
      gg1 <- gg1 + xlim(c(0,max(max_x1,max_x2, na.rm=TRUE)))
      gg2 <- gg2 + xlim(c(0,max(max_x1,max_x2, na.rm=TRUE)))

      #set size of y plots to be dependent on num cats
      #y axis has cats, this actually gets max of y axis, e.g. for 6 cats is 6.6
      max_y1 <- max(ggplot_build(gg1)$layout$panel_params[[1]]$y$continuous_range)
      max_y2 <- max(ggplot_build(gg2)$layout$panel_params[[1]]$y$continuous_range)

      #setting heights to num cats makes bar widths constant between cats
      gg1 / gg2 + plot_layout(heights=c(max_y1, max_y2)) #patchwork



```

## MoH facility attributes

```{r moh_facility_type_category, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=5}

# Facility_type_category
ggken <- facility_types("kenya", 
                        datasource=dfmoh,
                        type_column = 'Facility_type_category',
                        lonlat_columns = NULL,
                        #plot=FALSE,
                        plot_title = "MoH data : Frequency of facility type categories (no coords)")

#ggken  + plot_layout(ncol=2,widths=c(1,0.1))
```


```{r moh_keph_level, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=4}

# Keph level, defined facility levels
ggken <- facility_types("kenya", 
                        datasource=dfmoh,
                        type_column = "Keph level",
                        lonlat_columns = NULL,
                        plot_title = "MoH data : Keph level (no coords)")
#plot(ggken)
```


```{r moh_owner_type, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=3}

# Owner type = public/private
ggken <- facility_types("kenya", 
                        datasource=dfmoh,
                        type_column = "Owner type",
                        lonlat_columns = NULL,                        
                        plot_title = "MoH data : Owner type (no coords)")
#plot(ggken)


#as.data.frame(table(dfmoh$`Owner type`))
# 1       Faith Based Organization 1035
# 2             Ministry of Health 5875
# 3 Non-Governmental Organizations  366
# 4               Private Practice 5102


#number of beds - but the default plot doesn't cope well with the numeric data
# ggken <- facility_types("kenya", 
#                         datasource=dfmoh,
#                         type_column = 'Beds',
#                         lonlat_columns = NULL,                        
#                         plot_title = "MoH data : Number of beds (no coords)")
# plot(ggken)



```



```{r owner-wigiske, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=6}


#22 owners
owners_wigiske <- unique(sfwigiske[['Owner']])

ggken <- facility_types("kenya", 
                        datasource=sfwigiske,
                        type_column = "Owner",
                        plot_title = "WiGISKe data : Owner")


```

## All attribute columns in WiGISKe & MoH data 

```{r attribute columns, echo=FALSE, warning=FALSE}

#WiGISKe
names(sfwigiske)

#MoH
names(dfmoh)

```


```{r map_moh_kenya, eval=FALSE, include=FALSE}

#eval=FALSE because no coordinates

# just view MoH data
sfmoh <- afrihealthsites("kenya", datasource=file_moh,
                             type_column = 'Facility_type_category',
                             label_column = 'Facility Name',
                             lonlat_columns = c("Longitude", "Latitude"))

#testing filter
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter='Dispensary')
#filter with two types
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter=c('Dispensary','Health Center'))

#trying to remove one from selection doesn't work
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter='-Dispensary')
#sfmoh <- afrihealthsites("kenya", datasource=file_moh, type_filter='!Dispensary')

# plot moh vs kemri on a map
# this does work now but only because the defaults are set up for kenya
# todo sort code so will work for other countries
compare_hs_sources("kenya", datasources=c('who',file_moh),
                   type_column = 'Facility Type',
                   label_column = 'Facility Name')
                   #lonlat_columns = c("Longitude", "Latitude"))

```

# Number of facilities per ward in MoH data
```{r moh-per-ward, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=6}

library(dplyr)

dfcountbyward <- dfmoh %>% group_by(Ward) %>% summarise(count=n())

ggplot(dfcountbyward, aes(x=count)) + geom_bar()

```


```{r moh_type_and_owner_type, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=4}

# bar chart with columns for 'Facility_type_category', coloured by "Owner type"

#note 2 spaces after health
toexclude <- c("Primary health  care services", "None")
dfmoh2 <- dfmoh[-which(dfmoh$Facility_type_category %in% toexclude ),]

gg <- ggplot2::ggplot(dfmoh2, aes(y = Facility_type_category, fill = `Owner type`))

gg <- gg + geom_bar(show.legend=TRUE) +
      labs(title="MoH data - ownership by facility type") +
      theme_minimal() 
      
gg

```

```{r moh_kephlevel_and_owner_type, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=11, fig.height=4}


gg <- ggplot2::ggplot(dfmoh, aes(y = `Keph level`, fill = `Owner type`))

gg <- gg + geom_bar(show.legend=TRUE) +
      labs(title="MoH data - ownership by Keph level") +
      theme_minimal() 
      
gg

```


