---
title: "worldpop-ages"
output: html_document
author: andy south
---
  
Look at worldpop popn estimates by country and age ranges  

https://github.com/wpgp/wopr

demo: 
https://github.com/wpgp/wopr/blob/master/demo/wopr_demo.R

https://apps.worldpop.org/woprVision/

Vignette explaining how to access worldpop API directly :
https://github.com/wpgp/wopr/blob/master/vignettes/woprAPI.Rmd

Also different R package to access the worldpop API :
https://github.com/wpgp/wpCPR

DRC data in wopr - but popn is just for a few selected provinces in the east (in both v1 & 2) :
https://wopr.worldpop.org/?COD/

In contrast this does contain data for whole country, but not sure how to get at via R :
there do seem to be older worldpop data on age profiles for all of DRC
they are available as one tiff file per age/sex group
https://www.worldpop.org/geodata/summary?id=50153

TODO 
check out these packages :

https://github.com/TheDavidChen/FemaleWorldPop
seems to allow age & sex extraction of worldpop data

https://github.com/wpgp/wpgpDownloadR
and maybe others from here : https://www.worldpop.org/sdi/plugins



Note that main purpose of wopr seems to be to provide aggregated popn estimates at points or in polygons.

We could test out getting popn estimates by admin regions.

Note that the raster map displayed in the wopr vision app consists of tiles accessed from worldpop API.

Key functions :

woprize : Query WOPR to get population totals and confidence intervals for sf polygon or point features.
getPop : Query WOPR using an sf point or polygon to get population estimates as a vector of samples from the Bayesian posterior prediction for that location.

agesex_select	
character vector. Age-sex groups (e.g. "m0", "m1", "m5").
default is :
c(paste0("m", c(0, 1, seq(5, 80, 5))), paste0("f", c(0, 1, seq(5, 80,5))))
which gives :
"m0"  "m1"  "m5"  "m10" "m15" "m20" "m25" "m30" "m35" "m40" "m45" "m50" "m55" "m60" "m65" "m70" "m75" "m80" "f0" "f1"  "f5"  "f10" "f15" "f20" "f25" "f30" "f35" "f40" "f45" "f50" "f55" "f60" "f65" "f70" "f75" "f80"

So to get over 60s 
sexes <- c("m", "f")
ages <- seq(60, 80, 5)
c(paste0(rep(sexes,each=length(ages)),ages))

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

#library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(purrr) # for pluck
library(sf)
library(tmap)
library(lubridate)
library(afriadmin)

#remotes::install_github('wpgp/wopr')
library(wopr)

#remotes::install_github("wpgp/wpgpDownloadR")
library(wpgpDownloadR)

#using new path specification avoiding double slashes
#folder <- r"(C:\Dropbox\_afrimapr\data\covid-hera-hdx\)"

```


```{r, worldpop-wpgpDownloadR, eval=TRUE, warning=FALSE}

#remotes::install_github("wpgp/wpgpDownloadR")
#library(wpgpDownloadR)

#see also for code on extracting worldpop data e.g. by admin areas 
#https://github.com/TheDavidChen/FemaleWorldPop/blob/master/2_PopExtraction.R

#this code aggregates dataframe by age after earlier having calculated numbers per admin units
#https://github.com/TheDavidChen/FemaleWorldPop/blob/master/4_ageAggregation.R

dfdatasets <- wpgpListCountryDatasets(ISO3="COD") #cod = drc

#298 rows

#e.g.
#agesex_m_0_2020   Estimated 0-12 month old male per grid-cell  in 2020
#agesex_m_1_2020   Estimated 1-4 year old male per grid-cell  in 2020
#agesex_m_80_2020  Estimated 80 year old male per grid-cell  in 2020

#returns path of downloaded filr
datapath80 <- wpgpGetCountryDataset(ISO3 = "COD",
                            covariate = "agesex_m_80_2020" )
                            #    destDir ="G:\\WorldPop_Data")
#60
datapath60 <- wpgpGetCountryDataset(ISO3 = "COD",
                            covariate = "agesex_m_60_2020" )

library(raster)
rast80 <- raster::raster(datapath80)
rast60 <- raster::raster(datapath60)

plot(rast80)

dim(rast80)
#[1] 22610 22918 

mapview::mapview(rast80)
mapview::mapview(rast60) #is max really ~ 6 per 100m2 for male 60-65


#TODO we could write code to calculate estimated numbers of people by age & sex (e.g. over 60s)
#probably best to summarise by admin units first (e.g. health catchments and/or geoboundaries)
#to make calculations and data amounts more tractable

#beware these 100m resolution files are big, ~ 20k*20k cells, up to 1GB
#therefore good reason to convert to polygons

year <- 2020
age_start <- 60
age_end <- 80
layer <- 0
downloaded <- TRUE
destDir <- "C:\\Dropbox\\_afrimapr\\data\\worldpop"
countryiso <- "cod"

for( sex in c('f','m'))
{
  age <- age_start
  
  # Stop when the age exceeds the specified age_end
  while (age <= age_end) {

    # Create the link
    cov_name <- paste0("agesex_", sex, "_", age, "_", year)

    # save covariate names
    #cov_names <- rbind(cov_names, c(year, age, cov_name))
    
    
    #download data if not already
    if (!downloaded)
    {
      #takes ~ a few mins to download each layer
      datapath <- wpgpGetCountryDataset(ISO3 = "COD",
                                        covariate = cov_name,
                                        destDir = destDir)      
    } else
    {
      datapath <- paste0(destDir,"\\",countryiso,"_",cov_name,".tif")
    }
    
    
    rast <- raster::raster(datapath)
    
    # is it useful to add these to a raster stack
    # or to try to identify individual objects by name
    layer <- layer + 1
    
    if (layer==1) rstack <- raster::stack(rast) 
    else rstack <- raster::addLayer(rstack, rast)
    
    # ages available in intervals of 5
    age <- age + 5
  } 
  
}

# Sum all layers (ages & sexes) in the raster stack
rast_sum <- sum(rstack) #takes ~ 12 mins

mapview::mapview(rast_sum)

#or r <- calc(s, fun=sum). 
#calc should be faster when using complex formulas on large datasets & is possible to set an output filename and file type preferences.

#takes a few minutes
writeRaster(rast_sum, filename=paste0(destDir,"\\",countryiso,"_",age_start,"to",age_end,".tif"))
#"C:\\Dropbox\\_afrimapr\\data\\worldpop\\cod_60to80.tif"

```


```{r, worldpop-wopr-demo-egs, eval=TRUE, warning=FALSE}


##---- population estimates for a single polygon ----##

# get population total
N <- getPop(feature=wopr_polys[1,],
            country='NGA',
            version='v1.2')

# summarize population total
summaryPop(N, confidence=0.95, tails=2, belowthresh=1e5, abovethresh=2e5)
hist(N)

##---- population estimates for multiple features ----##
# example from demo, takes ~ 30 seconds to run, accesses worldpop API

#andy to get over 60s 
sexes <- c("m", "f")
ages <- seq(60, 80, 5)
over60 <- c(paste0(rep(sexes,each=length(ages)),ages))

# get population totals
totals <- woprize(features=wopr_polys,
                  country='NGA',
                  version='v1.2',
                  #my age selction >=60
                  agesex_select=over60,
                  confidence=0.95,
                  tails=2,
                  belowthresh=1e5,
                  abovethresh=2e5
                  )

#to view df without geometry
sf::st_drop_geometry(totals)

tmap::tm_shape(totals) + tmap::tm_fill('mean', palette='Reds', legend.reverse=T)



#mapview::mapview(sfareas, zcol='edite_par', label=paste(sfareas$AS,"in",sfareas$ZS))



```