---
title: "worldpop-ages"
output: html_document
author: andy south
---
  
Look at worldpop popn estimates by country and age ranges  

https://github.com/wpgp/wopr

demo: 
https://github.com/wpgp/wopr/blob/master/demo/wopr_demo.R

https://apps.worldpop.org/woprVision/

Vignette explaining how to access worldpop API directly :
https://github.com/wpgp/wopr/blob/master/vignettes/woprAPI.Rmd

Also different R package to access the worldpop API :
https://github.com/wpgp/wpCPR

DRC data in wopr - but popn is just for a few selected provinces in the east (in both v1 & 2) :
https://wopr.worldpop.org/?COD/

In contrast this does contain data for whole country, but not sure how to get at via R :
there do seem to be older worldpop data on age profiles for all of DRC
they are available as one tiff file per age/sex group
https://www.worldpop.org/geodata/summary?id=50153

TODO check out this pacakge :
https://github.com/wpgp/wpgpDownloadR
and maybe others from here : https://www.worldpop.org/sdi/plugins


Note that main purpose of wopr seems to be to provide aggregated popn estimates at points or in polygons.

We could test out getting popn estimates by admin regions.

Note that the raster map displayed in the wopr vision app consists of tiles accessed from worldpop API.

Key functions :

woprize : Query WOPR to get population totals and confidence intervals for sf polygon or point features.
getPop : Query WOPR using an sf point or polygon to get population estimates as a vector of samples from the Bayesian posterior prediction for that location.

agesex_select	
character vector. Age-sex groups (e.g. "m0", "m1", "m5").
default is :
c(paste0("m", c(0, 1, seq(5, 80, 5))), paste0("f", c(0, 1, seq(5, 80,5))))
which gives :
"m0"  "m1"  "m5"  "m10" "m15" "m20" "m25" "m30" "m35" "m40" "m45" "m50" "m55" "m60" "m65" "m70" "m75" "m80" "f0" "f1"  "f5"  "f10" "f15" "f20" "f25" "f30" "f35" "f40" "f45" "f50" "f55" "f60" "f65" "f70" "f75" "f80"

So to get over 60s 
sexes <- c("m", "f")
ages <- seq(60, 80, 5)
c(paste0(rep(sexes,each=length(ages)),ages))

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

#library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(purrr) # for pluck
library(sf)
library(tmap)
library(lubridate)
library(afriadmin)


#remotes::install_github('wpgp/wopr')
library(wopr)

#using new path specification avoiding double slashes
#folder <- r"(C:\Dropbox\_afrimapr\data\covid-hera-hdx\)"

```


```{r, worldpop-wopr-demo-egs, eval=TRUE, warning=FALSE}


##---- population estimates for a single polygon ----##

# get population total
N <- getPop(feature=wopr_polys[1,],
            country='NGA',
            version='v1.2')

# summarize population total
summaryPop(N, confidence=0.95, tails=2, belowthresh=1e5, abovethresh=2e5)
hist(N)

##---- population estimates for multiple features ----##
# example from demo, takes ~ 30 seconds to run, accesses worldpop API

#andy to get over 60s 
sexes <- c("m", "f")
ages <- seq(60, 80, 5)
over60 <- c(paste0(rep(sexes,each=length(ages)),ages))

# get population totals
totals <- woprize(features=wopr_polys,
                  country='NGA',
                  version='v1.2',
                  #my age selction >=60
                  agesex_select=over60,
                  confidence=0.95,
                  tails=2,
                  belowthresh=1e5,
                  abovethresh=2e5
                  )

#to view df without geometry
sf::st_drop_geometry(totals)

tmap::tm_shape(totals) + tmap::tm_fill('mean', palette='Reds', legend.reverse=T)



#mapview::mapview(sfareas, zcol='edite_par', label=paste(sfareas$AS,"in",sfareas$ZS))



```