---
title: "openstreetmap-africa-explore"
output: html_document
---
  
## starting to look at accessing useful OSM data for Africa

ISSUE :
How best to extract data for particular countries

Or first example to extract motorways or similar for the whole continent to have an example lines dataset for training

osmdata package seems best.
https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html
https://dominicroye.github.io/en/2018/accessing-openstreetmap-data-with-r/

opq_string() gives overpass query to test at : https://overpass-turbo.eu/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!require("osmdata")) install.packages("osmdata")

library(dplyr)
library(ggplot2)
library(readr)
library(osmdata)

```

```{r, osmdata-examples, eval=TRUE, warning=FALSE}

#starting example from osmdata vignette
sf1 <- opq(bbox = 'greater london uk') %>%  
    add_osm_feature(key = 'highway', value = 'motorway')  %>%
    #opq_string()
    osmdata_sf()
plot(sf::st_geometry(sf1$osm_lines))
mapview(sf1$osm_lines)

# trimming data by polygon bbox
# bb <- getbb ('london uk', format_out = 'polygon')
# sf1 <- opq(bbox = bb) %>%
#     add_osm_feature(key = 'highway', value = 'motorway') %>%
#     osmdata_sf () %>%
#     trim_osmdata (bb)

```


```{r, africa-motorways, eval=TRUE, warning=FALSE}

# testing getting bboxes for africa
bbkenya <- getbb("kenya", featuretype = 'country')
# x 33.90969 41.90675
# y -4.89952  4.62000
bbalgeria <- getbb("algeria", featuretype = 'country')

bbafrica <- getbb("africa", featuretype = 'continent')
# x  -7.242188 42.75781
# y -13.497566 36.50243

#note getbb(format_out='sf_polygon') #to be able to plot in sf
sfbb <- getbb("kenya", featuretype='country', format_out='sf_polygon')

sfbb <- getbb("africa", featuretype = 'continent', format_out='sf_polygon')
#but this gave 2 polygons one in Venezuela one in Kansas
#mapview(sfbb[1,])
#mapview(sfbb[2,])

#dataframe returns 10 results, the first is the continent
dfbb <- getbb("africa", featuretype = 'continent', format_out='data.frame')
dfbb[1]
# but seems I can't get a polygon for it
# dfbb[1,'boundingbox']
# "-13.4975662" "36.5024338"  "-7.2421878"  "42.7578122" 

# coords seem the same as from getbb with default feature type
# BUT seems to me not to include full africa extent
# eg not far west of senegal
#getbb("senegal", featuretype = 'country')
# x -17.78624 -11.34594
# y  12.23728  16.69197


#NOTE features are added as an AND operation
#to achieve OR have to do separate queries & combine with c(a,b)

#timeout in secs, here 3 mins

sf1 <- opq(timeout=180, bbox = bbkenya) %>%  
    #add_osm_feature(key = 'highway', value = 'motorway')  %>%
    #add_osm_feature(key = 'highway', value = 'trunk')  %>% 
    add_osm_feature(key = 'highway', value = 'primary')  %>%  
    #opq_string() #query to test at https://overpass-turbo.eu/
    osmdata_sf()
plot(sf::st_geometry(sf1$osm_lines))

mapview(sf1$osm_lines)

# downloading motorways for algeria worked but seemingly none in Kenya
sfmw_alg <- opq(timeout=900, bbox = bbalgeria) %>%  
    add_osm_feature(key = 'highway', value = 'motorway')  %>%
    #add_osm_feature(key = 'highway', value = 'trunk')  %>% 
    #add_osm_feature(key = 'highway', value = 'primary')  %>%  
    osmdata_sf()
plot(sf::st_geometry(sfmw$osm_lines))

# try to download all motorways for Africa
# but Africa bbox looks like it misses some of W
# failed a few times but did eventually get there 
# lines have 10458 rows & 180 columns so not really good teaching data !
sfmw <- opq(timeout=900, bbox = bbafrica) %>%  
    add_osm_feature(key = 'highway', value = 'motorway')  %>%
    #add_osm_feature(key = 'highway', value = 'trunk')  %>% 
    #add_osm_feature(key = 'highway', value = 'primary')  %>%  
    osmdata_sf()
plot(sf::st_geometry(sfmw$osm_lines))

# different errors each time
# Request failed [504]. Retrying in 3.5 seconds...
# Error in read_xml.raw(charToRaw(enc2utf8(x)), "UTF-8", ..., as_html = as_html,  : 
#   Space required after the Public Identifier [65]

# Error in read_xml.raw(charToRaw(enc2utf8(x)), "UTF-8", ..., as_html = as_html,  : 
#   Opening and ending tag mismatch: hr line 5 and body [76]

# trimming data by polygon bbox
# bb <- getbb ('london uk', format_out = 'polygon')
# sf1 <- opq(bbox = bb) %>%
#     add_osm_feature(key = 'highway', value = 'motorway') %>%
#     osmdata_sf () %>%
#     trim_osmdata (bb)


```

# trying global roads inventory project
# https://www.globio.info/download-grip-dataset
# GRIP mainly aimed at providing a roads dataset that is easily usable for scientific global environmental and biodiversity modelling projects. The dataset is not suitable for navigation. GRIP4 is provided under an Open Data Commons Open Database License (ODbL) and is free to use.
# https://dataportaal.pbl.nl/downloads/GRIP4/GRIP4_Region3_vector_shp.zip

# looks good in QGIS
# road types highways and primary roads
# gp_rtp 1 & 2

```{r, africa-roads-grip, eval=TRUE, warning=FALSE}

filepath <- r"(C:\Dropbox\_afrimapr\road-data\GRIP4_Region3_vector_shp\GRIP4_Region3.shp)"
  
sfroads <- sf::st_read(filepath)  #>1 million rows !

sfhways <- dplyr::filter(sfroads, GP_RTP==1) #5069 rows mostly around edges

#trying kmz of transcontinental network
filepath <- r"(C:\Dropbox\_afrimapr\road-data\Trans-African Highway Network.kml)"
  
sfcontinental <- sf::st_read(filepath)


```



```{r, africa-osm-airports, eval=TRUE, warning=FALSE}

# aeroway=aerodrome

airp <- opq(timeout=900, bbox = bbafrica) %>%  
    add_osm_feature(key = 'aeroway', value = 'aerodrome')  %>%
    osmdata_sf()

#choose nodes because interested in points 33677 points
plot(sf::st_geometry(airp$osm_points))

# does show that there is problem with the africa bbox, misses parts of w & south
# and the nodes are lots of points per airport at least in some cases

sfairp <- airp$osm_points %>% 
          dplyr::filter(!is.na(name)) %>% 
          dplyr::filter(!(is.na(iata) | is.na(icao))) %>% 
          dplyr::filter(!is.na(aeroway))
          
#down to 1401 with a name (lots of dodgy encoding)
#1467 not na aeroway
#1398 both
#166 columns icao & iata seem to be codes that are in a subset down to 166

#still plenty dodgy encoding e.g. مطار الوجه المحلي

mapview::mapview(sfairp, zcol='name', legend=FALSE)

# test for rwanda
bbrwanda <- getbb("rwanda", featuretype = 'country')
airp_rwa <- opq(timeout=900, bbox = bbrwanda) %>%  
    add_osm_feature(key = 'aeroway', value = 'aerodrome')  %>%
    osmdata_sf()

sfairp_rwa <- airp_rwa$osm_points

#458 rows

# only few have anything in name field
# these cut to 7 but no Kigali 
# maybe I could edit in OSM ??
sfairp_rwa2 <- dplyr::filter(sfairp_rwa, !is.na(name))
sfairp_rwa2 <- dplyr::filter(sfairp_rwa, !is.na(aeroway))

mapview::mapview(sfairp_rwa)
mapview::mapview(sfairp_rwa2)

```


```{r, africa-railways, eval=TRUE, warning=FALSE}


bbafrica <- getbb("africa", featuretype = 'continent')


# try to download all motorways for Africa
# but Africa bbox looks like it misses some of W
# failed a few times but did eventually get there 
# lines have 10458 rows & 180 columns so not really good teaching data !
sfrw <- opq(timeout=900, bbox = bbafrica) %>%  
    add_osm_feature(key = 'railway', value = 'rail')  %>%
    osmdata_sf()

plot(sf::st_geometry(sfrw$osm_lines))




```
