---
title: "join-to-admin-rwanda"
output: html_document
---
  
Experimenting with joining data to sub-national admin boundaries.

ISSUE :
There are usually difficulties in joining data to admin boundaries due to spelling, capitalisation and accents.

afrimapr wants to improve that.

First want to document some of the difficulties in the wild.

Also this can be part of upcoming training material.

Reference : 
Good outlines in here that we can borrow some parts from.
https://geocompr.robinlovelace.net/attr.html#vector-attribute-joining
Vignette on joins that goes further, still doesn't provide much advice on how to join when names don't match :
https://geocompr.github.io/geocompkg/articles/join.html
From R for datascience, some good diagrams but terminology a bit advanced. Short section on join problems.
https://r4ds.had.co.nz/relational-data.html#join-problems
"Check that your foreign keys match primary keys in another table. The best way to do this is with an anti_join(). It’s common for keys not to match because of data entry errors. Fixing these is often a lot of work."

"Be aware that simply checking the number of rows before and after the join is not sufficient to ensure that your join has gone smoothly. If you have an inner join with duplicate keys in both tables, you might get unlucky as the number of dropped rows might exactly equal the number of duplicated rows!"

fuzzyjoin package :
https://github.com/dgrtwo/fuzzyjoin

safejoin package looks good but in early development and not on CRAN :
could provide inspiration for advice to provide to new users 
https://github.com/moodymudskipper/safejoin
remotes::install_github("moodymudskipper/safejoin")


https://github.com/mmurtaghw/Census-Visualisation-Rwanda/blob/master/src/Vulnerability_Mapping.R

Rwanda, number of admin units
adm1 5    provinces  
adm2 30   districts
adm3 ~416 sectors
adm4    cells


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode)

# install.packages("remotes")

#remotes::install_github("afrimapr/afriadmin")
library(afriadmin) 
library(mapview)
library(knitr) #for kable
library(dplyr)
library(ggplot2)
library(fuzzyjoin)

remotes::install_gitlab("dickoa/rgeoboundaries")
library(rgeoboundaries)

```


```{r, eval=TRUE, warning=FALSE}

#getting metadata on boundaries available

metarwa <- rgeoboundaries::gb_metadata('rwanda')

metarwa1 <-rgeoboundaries::gb_metadata('rwanda', adm_lvl = 'adm1')

#rwanda has up to adm4

#getting admin boundaries using rgeoboundaries() could also use afriadmin that tries to make simpler
sfadm1 <- rgeoboundaries::geoboundaries('rwanda', adm_lvl="adm1", type="sscu")
sfadm2 <- rgeoboundaries::geoboundaries('rwanda', adm_lvl="adm2", type="sscu")
sfadm3 <- rgeoboundaries::geoboundaries('rwanda', adm_lvl="adm3", type="sscu")
sfadm4 <- rgeoboundaries::geoboundaries('rwanda', adm_lvl="adm4", type="sscu")

plot(sf::st_geometry(sfadm1))
plot(sf::st_geometry(sfadm2))
plot(sf::st_geometry(sfadm3))
plot(sf::st_geometry(sfadm4))

#mapview::mapview(sfadm3, zcol = "shapeName")
#mapview::mapview(sfadm4, zcol = "shapeName")

```

```{r, csv-data-download, eval=TRUE, warning=FALSE}

#download some data to try to join to admin boundaries
inurl <- "https://data.humdata.org/dataset/94b26815-cd91-4553-b64d-7a6f153400bc/resource/b07265ad-055e-4ea8-bb95-de3adb1ae0a1/download/rwa_adm3_pop_rphc2012.csv"

#dfpop_adm3 <- read.csv(inurl)

#annoyingly numeric fields get read in as chars with thousands commas
 # $ ADM3_EN   : chr  "Gitega" "Kanyinya" "Kigali" "Kimisagara" ...
 # $ ADM3_PCODE: chr  "RW1101" "RW1102" "RW1103" "RW1104" ...
 # $ ADM3ALT1EN: chr  "" "" "" "" ...
 # $ Rwanda_T  : chr  "28,728" "21,859" "30,023" "46,753" ...
 # $ Rwanda_M  : chr  "14,989" "10,777" "15,375" "24,451" ...

# use readr to get around commas in numbers problem
library(readr)
#specify which columns are characters and numbers (number copes with commas for thousands)
#I counted by hand that 1st 14 columns are character variables, and the 9 after are numbers
dfpop_adm3 <- readr::read_csv(inurl, col_types = c(rep(col_character(),14),rep(col_number(),9)))

# STARTING TO ARRANGE WORKFLOW FOR LOOKING AT THE DATA PRIOR TO ATTEMPTING JOIN
# could create a decision tree or flow chart

nrow(dfpop_adm3)
#416 rows

#ADM3_EN contains admin3 names
dfpop_adm3$ADM3_EN

 #  [1] "Gitega"       "Kanyinya"     "Kigali"       "Kimisagara"   "Mageregere"   "Muhima"       "Nyakabanda"
 #  [8] "Nyamirambo"   "Nyarugenge"   "Rwezamenyo"   "Bumbogo"      "Gatsata"      "Gikomero"     "Gisozi"    
 # [15] "Jabana"       "Jali"         "Kacyiru"      "Kimihurura"   "Kimironko"    "Kinyinya"     "Ndera"   

#visual inspection shows that in Title case with no accents 

#check if all names are unique
length(unique(dfpop_adm3$ADM3_EN))
#379 #so most, but not all are unique

#I wonder if the ones that are not unique represent the same name in different areas ?

# BOUNDARIES FILE

#what column contains the names of the admin areas ?
#str shows us structure of the object
#promising that it shows 416 rows which is the same as our data file
str(sfadm3)
# Classes ‘sf’ and 'data.frame':	416 obs. of  6 variables:
#  $ shapeName : chr  "Gitega" "Kanyinya" "Kigali" "Kimisagara" ...
#  $ shapeISO  : chr  "None" "None" "None" "None" ...
#  $ shapeID   : chr  "RWA-ADM3-3_0_0-B1" "RWA-ADM3-3_0_0-B2" "RWA-ADM3-3_0_0-B3" "RWA-ADM3-3_0_0-B4" ...
#  $ shapeGroup: chr  "RWA" "RWA" "RWA" "RWA" ...
#  $ shapeType : chr  "ADM3" "ADM3" "ADM3" "ADM3" ...
#  $ geometry  :sfc_MULTIPOLYGON of length 416; first list element: List of 1

# shapeName is the column we want because it contains the differing names of each admin area

# does the shapeName column also contain non-unique names like the population data file does ?
length(unique(sfadm3$shapeName))
# yes, also 379

# Aha seems to show that no names are unique to just one of the datasets
setdiff(dfpop_adm3$ADM3_EN, sfadm3$shapeName )
#character(0)

#So tricky that we want to look at the repeated names 
#the pop data file has ADM2_EN names too
dfpop3_dups <- dfpop_adm3[ duplicated(dfpop_adm3$ADM3_EN), ]
#this does show some as suspected where same adm3 names are in diff adm2 regions (e.g. Kageyo)
#BUT also some seem not to be duplicates (e.g. Busasamana)
sort(dfpop3_dups$ADM3_EN)
#  [1] "Busasamana" "Cyanika"    "Kageyo"     "Kageyo"     "Karama"     "Karama"     "Kibirizi"   "Kigarama"  
#  [9] "Kigoma"     "Kinazi"     "Kinihira"   "Mbazi"      "Muganza"    "Muganza"    "Mukarange"  "Muko"      
# [17] "Mukura"     "Murama"     "Murambi"    "Murambi"    "Murundi"    "Musha"      "Nemba"      "Ngoma"     
# [25] "Ngoma"      "Nyagisozi"  "Nyamiyaga"  "Nyange"     "Nyarugenge" "Remera"     "Remera"     "Remera"    
# [33] "Rugarama"   "Ruhango"    "Rukomo"     "Rwimbogo"   "Tumba"  

#does give 2 results
busasamana <- dfpop_adm3[ dfpop_adm3$ADM3_EN == "Busasamana", ]
#issue because duplicated does this : determines which elements of a vector or data frame are duplicates of elements with smaller subscripts

#must be better tidyverse solution
#distinct(dfpop_adm3,ADM3_EN) # this does same as unique

dfpop3_dups <- 
  dfpop_adm3 %>%
  #group by the name column
  group_by(ADM3_EN) %>%
  #add a column with number of each name
  mutate(n=n()) %>%
  #filter names that appear more than once
  filter(n>1) 

#can do this for the boundaries too
sfadm3_dups <- 
  sfadm3 %>%
  #group by the name column
  group_by(shapeName) %>%
  #add a column with number of each name
  mutate(n=n()) %>%
  #filter names that appear more than once
  filter(n>1) 

# can I write a function to do that
# relying on dplyr or in base ?
# this keeps the returned object as sf if it was before
# if no duplicates, it returns df with nrow=0, I could get it to return NULL or 0
filter_duplicates <- function(df, column_name) 
{
  dfdup <- df %>%
    #group by the name column
    group_by(.data[[column_name]]) %>%
    #add a column with number of each name
    mutate(n=n()) %>%
    #filter names that appear more than once
    filter(n>1) 
  
  if (nrow(dfdup)==0) {
    warning("no duplicates")
    dfdup <- NULL
  }
  
  return(dfdup)
}


#plot just duplicates
plot(sf::st_geometry(sfadm3_dups))
mapview::mapview(sfadm3_dups)

#colour by shapename - slightly too many to see pairs clearly
mapview::mapview(sfadm3_dups, zcol='shapeName')

#Probably want to join with multiple keys (shared columns) like below
#BUT the regoboundaries file doesn't have the adm2 names that are needed
# sfjoined <- left_join(sfadm3, dfpop_adm3, 
#                       by = c("ADM2_EN", "shapeName" = "ADM3_EN"))


#try loading the boundary data from hdx
#https://data.humdata.org/dataset/2768bdfd-6486-4963-8e3d-e63149478eb4/resource/176eaace-708c-46b3-94e9-d5ba91cca08c/download/rwa_adm_2006_nisr_wgs1984_20181002_shp.zip

# library(rhdx)
# 
# set_rhdx_config(hdx_site = "prod")
# list_of_ds <- search_datasets("rwanda boundaries", rows = 2)
# list_of_ds

#admin boundaries do come up first
# [[1]]
# <HDX Dataset> 2768bdfd-6486-4963-8e3d-e63149478eb4 
#   Title: Rwanda administrative levels 0-4 boundaries
#   Name: rwanda-administrative-boundaries-level-1-4
#   Date: 09/27/2018
#   Tags (up to 5): administrative divisions, common operational dataset - cod, gazetteer, geodata
#   Locations (up to 5): rwa
#   Resources (up to 5): rwa_gazetteer_rphc2012.xlsx, rwa_adm_2006_NISR_WGS1984_20181002_SHP.zip

# each dataset consists of a number of resources
# (see https://github.com/dickoa/rhdx)

#ds <- list_of_ds[[1]]

# see what resources there are
#get_resources(ds)

# [[1]]
# <HDX Resource> 95b0f509-7bbb-44cb-a783-92c8208fdfea 
#   Name: rwa_gazetteer_rphc2012.xlsx
#   Description: Rwanda administrative level 0 (country), 1 (province), 2 (district), 3 (sector), and 4 (cell [incomplete]) gazetteer
#   Size: 167305
#   Format: XLSX
# 
# [[2]]
# <HDX Resource> 176eaace-708c-46b3-94e9-d5ba91cca08c 
#   Name: rwa_adm_2006_NISR_WGS1984_20181002_SHP.zip
#   Description: Rwanda administrative level 0 (country), 1 (province), 2 (district), 3 (sector), and 4 (cell [incomplete]) gazetteer
#   Size: 15484795
#   Format: ZIPPED SHAPEFILES

# we want the 2nd set, the zipped shapefiles

# first get information about the first resource
#rs <- get_resource(ds, 1)

# find which layers in file
#mlayers <- rhdx::get_resource_layers(rs, download_folder=getwd())

#arrg
#Error: This (spatial) data format is not yet supported - i think maybe because multiple layers in one zip

# then download the datafiles
#sfrwa <- read_resource(rs, download_folder = tempdir())


# instead download the zip by hand from HDX
# https://data.humdata.org/dataset/rwanda-administrative-boundaries-level-1-4

# choose the admin 3 filename, ending .shp
filename <- "rwa_adm3_2006_NISR_WGS1984_20181002.shp"
#double slashes are just to avoid potential confusion between windows and linux
folder <- "C:\\Dropbox\\_afrimapr\\hdx\\"

filename <- paste0(folder, filename)

# read data from file into sf object
sfrwa_adm3 <- sf::st_read(filename)

# check str(ucture) of new object
str(sfrwa_adm3)
# Classes ‘sf’ and 'data.frame':	416 obs. of  15 variables:
#  $ ADM0_EN   : chr  "Rwanda" "Rwanda" "Rwanda" "Rwanda" ...
#  $ ADM0_FR   : chr  "Rwanda (le)" "Rwanda (le)" "Rwanda (le)" "Rwanda (le)" ...
#  $ ADM0_RW   : chr  "U Rwanda" "U Rwanda" "U Rwanda" "U Rwanda" ...
#  $ ADM0_SW   : chr  "Rwanda" "Rwanda" "Rwanda" "Rwanda" ...
#  $ ADM0_PCODE: chr  "RW" "RW" "RW" "RW" ...
#  $ ADM1_EN   : chr  "Kigali City" "Kigali City" "Kigali City" "Kigali City" ...
#  $ ADM1_FR   : chr  "Ville de Kigali" "Ville de Kigali" "Ville de Kigali" "Ville de Kigali" ...
#  $ ADM1_RW   : chr  "Umujyi wa Kigali" "Umujyi wa Kigali" "Umujyi wa Kigali" "Umujyi wa Kigali" ...
#  $ ADM1_PCODE: chr  "RW1" "RW1" "RW1" "RW1" ...
#  $ ADM2_EN   : chr  "Nyarugenge" "Nyarugenge" "Nyarugenge" "Nyarugenge" ...
#  $ ADM2_PCODE: chr  "RW11" "RW11" "RW11" "RW11" ...
#  $ ADM3_EN   : chr  "Gitega" "Kanyinya" "Kigali" "Kimisagara" ...
#  $ ADM3_PCODE: chr  "RW1101" "RW1102" "RW1103" "RW1104" ...
#  $ ADM3ALT1EN: chr  NA NA NA NA ...
#  $ geometry  :sfc_MULTIPOLYGON of length 416; first list element: List of 1
#   ..$ :List of 1
#   .. ..$ : num [1:464, 1:2] 30 30 30 30 30 ...
#   ..- attr(*, "class")= chr [1:3] "XY" "MULTIPOLYGON" "sfg"
#  - attr(*, "sf_column")= chr "geometry"
#  - attr(*, "agr")= Factor w/ 3 levels "constant","aggregate",..: NA NA NA NA NA NA NA NA NA NA ...
#   ..- attr(*, "names")= chr [1:14] "ADM0_EN" "ADM0_FR" "ADM0_RW" "ADM0_SW" ...

# good that it has 416 obs (same as the data file that we want to join to it)
# and that it has columns containing names of the admin3 regions, and the admin2 regions that they exists in
# (allowing us to differentiate between admin3 regions with the same names)
# ADM3_EN ADM2_EN

# these are the same column names as in our population datafile 
#checking on matches of names, this time produces no non matches
non_matches <- str_subset(sfrwa_adm3$ADM3_EN, dfpop_adm3$ADM3_EN, negate = TRUE)

#making the join based on the two columns
sfjoined <- dplyr::left_join(sfrwa_adm3, dfpop_adm3,
                       by = c("ADM2_EN", "ADM3_EN"))

# if columns had been named differently in each object
# left_join(sf, df, by = c("name1_in_obj1"="name1_in_obj2", "name2_in_obj1"="name2_in_obj2"))

names(sfjoined)
# can see the population data (e.g. Rwanda_T, Rwanda_M, Urban_T) have been added on to the geometry column

#  [1] "ADM0_EN.x"    "ADM0_FR.x"    "ADM0_RW.x"    "ADM0_SW.x"    "ADM0_PCODE.x" "ADM1_EN.x"    "ADM1_FR.x"   
#  [8] "ADM1_RW.x"    "ADM1_PCODE.x" "ADM2_EN"      "ADM2_PCODE.x" "ADM3_EN"      "ADM3_PCODE.x" "ADM3ALT1EN.x"
# [15] "ADM0_EN.y"    "ADM0_FR.y"    "ADM0_RW.y"    "ADM0_SW.y"    "ADM0_PCODE.y" "ADM1_EN.y"    "ADM1_FR.y"   
# [22] "ADM1_RW.y"    "ADM1_PCODE.y" "ADM2_PCODE.y" "ADM3_PCODE.y" "ADM3ALT1EN.y" "Rwanda_T"     "Rwanda_M"    
# [29] "Rwanda_F"     "Urban_T"      "Urban_M"      "Urban_F"      "Rural_T"      "Rural_M"      "Rural_F"     
# [36] "geometry"

#duplicate columns in the two objects are give the suffixes .x & .y 

# see below for what would have happened if we had joined just based on the admin3 column

# view total popn. in an interactive map
mapview::mapview(sfjoined, zcol='Rwanda_T')

# view total popn. in a static map (using sf)
plot( sfjoined["Rwanda_T"])

# view gender ratio F/M
sfjoined$f_div_m <- sfjoined$Rwanda_F/sfjoined$Rwanda_M
mapview::mapview(sfjoined, zcol='f_div_m')

# compare gender ratio and urban rural ratio side by side in static maps
sfjoined$rur_div_urb <- sfjoined$Rural_T/sfjoined$Urban_T

# view total popn. in a static map (using sf)
plot( sfjoined[c('f_div_m','rur_div_urb')])

# note, these data also both contain a unique column ADM3_PCODE which could have been used to join them
# but not all boundary and data files contain pcodes


# what would have happened if we had joined just based on the admin3 column ?
sfjoin_justadm3 <- dplyr::left_join(sfrwa_adm3, dfpop_adm3, by = "ADM3_EN")

#this looks fine, no clue that anything is wrong
mapview::mapview(sfjoin_justadm3, zcol='Rwanda_T')

#str(sfjoin_justadm3)
#Classes ‘sf’ and 'data.frame':	506 obs. of  37 variables:
#506 rows, where we only had 416 rows before

#TODO, how is it possible to give a warning that something unexpected has happened ?
# i.e. extra rows have been added due to the duplicates
# as Hadley points out in r4ds checking nrows is not enough
# because could get situation where some are duplicated and others not matched at all

# again we can check whether there are duplicates in the names using my function from above
dups_after_join <- filter_duplicates(sfjoin_justadm3, "ADM3_EN")

# but remember that duplicates is possibly a rarer problem than names that don't match

# can I find some Rwanda data on HDX (or Rwandan Gov) that demos the non matching issue ?
# http://www.devinfo.statistics.gov.rw/di7web/libraries/aspx/Catalog.aspx


# from HDX Global Multidimensional Poverty Index (MPI)
# it only has for 5 regions, not very interesting
# https://data.humdata.org/dataset/rwanda-mpi
# https://data.humdata.org/dataset/06b01ee1-5691-4c25-af69-e50bde0f5b57/resource/250bd322-098b-4761-a08a-ec2d77df54f7/download/rwa-subnational-results-mpi-2020.xlsx
# mpiurl <- "https://data.humdata.org/dataset/06b01ee1-5691-4c25-af69-e50bde0f5b57/resource/250bd322-098b-4761-a08a-ec2d77df54f7/download/rwa-subnational-results-mpi-2020.xlsx"
# # tried to read in excel file directly but not well formatted for machine reading
# library(readxl)
# library(httr)
# httr::GET(mpiurl, write_disk(tf <- tempfile(fileext = ".xlsx")))
# df <- readxl::read_excel(tf)


# http://www.devinfo.statistics.gov.rw/di7web/libraries/aspx/Catalog.aspx

filename <- "C:\\Users\\andy.south\\Google Drive\\afrimapr-gdrive\\training\\rwanda-training\\rwanda-data\\DevInfoRwanda_2013_en.csv"

df1 <- readr::read_csv(filename)

str(df1)
# tibble [30,788 x 9]
 # $ Indicator  : chr [1:30788] "1.02 Poverty gap ratio" "1.02 Poverty gap ratio" "1.02 Poverty gap ratio" 
 # $ Unit       : chr [1:30788] "Percent" "Percent" "Percent" "Percent" ...
 # $ Subgroup   : chr [1:30788] "Total" "Total" "Total" "Rural Access to Internet: Cyber Cafe" ...
 # $ Area       : chr [1:30788] "Rwanda" "Rwanda" "Rwanda" "Rwanda" ...
 # $ Area ID    : num [1:30788] 5 5 5 5 5 5 5 6 9 10 ...
 # $ Time Period: chr [1:30788] "2000-2001" "2005-2006" "2010-2011" "2012" ...
 # $ Source     : chr [1:30788] 
 # $ Data Value : num [1:30788] 24.4 22.3 14.8 1.1 0.7 1.2 58.9 22.7 65.5 62.3 ...
 # $ Footnotes  : logi [1:30788] NA NA NA NA NA NA ...


# after I registered I downloaded this 10% sample of the 2012 census
# not allowed to redistribute
# https://microdata.statistics.gov.rw/index.php/catalog/65/get_microdata
# (may also be some other data sources there)

# formatting and visualisation code here :
# https://github.com/mmurtaghw/Census-Visualisation-Rwanda


```


```{r, point-data-airports, eval=TRUE, warning=FALSE}

# airports from RNIS via hdx
# Rwanda Airports as published by Rwandan National Institute of Statistics
# https://data.humdata.org/dataset/rwanda-ports
# click to download & unzip

filename <- "C:\\Users\\andy.south\\Google Drive\\afrimapr-gdrive\\training\\rwanda-training\\rwanda-data\\RWA_Airports_NISR.shp"

# read in a shapefile (shp) with sf package
sfairports <- sf::st_read(filename)

# only 3 
str(sfairports)
# Classes ‘sf’ and 'data.frame':	3 obs. of  7 variables:
#  $ NAME      : chr  "KIGALI" "GISENYI" "KAMEMBE"
#  $ USE       : chr  "Joint Military/Civilian" "Other" "Other"
#  $ CLASS_NAME: chr  "RW45746" "RW21133" "RW43879"
#  $ FEATURE   : chr  "Airport/Airfield" "Airport/Airfield" "Airport/Airfield"
#  $ ICAO      : chr  "HRYR" "HRYG" "HRZA"
#  $ ELEVATION : chr  "1491" "1556" "1591"
#  $ geometry  :sfc_MULTIPOINT of length 3; first list element:  'XY' num [1, 1:2] 30.14 -1.97
mapview::mapview(sfairports)

# hdx also has one from OpenStreetMap
# https://data.humdata.org/dataset/hotosm_rwa_airports
# download the first one that is a points shapefile
filename <- "C:\\Users\\andy.south\\Google Drive\\afrimapr-gdrive\\training\\rwanda-training\\rwanda-data\\hotosm_rwa_airports_points.shp"

# read in a shapefile (shp) with sf package
sfosmairports <- sf::st_read(filename)

str(sfosmairports)
#6 airports of which 3 are helipads, but doesn't have Kigali
#Classes ‘sf’ and 'data.frame':	6 obs. of  12 variables:

mapview::mapview(sfosmairports)

# both files have a geometry column, this is what places the data on a map
# in both cases there are two numbers per row, the coordinates of a single point
# the data are classed differently one as POINT and one as MULTIPOINT but that doesn't matter for us now
# (MULTIPOINT allows for the potential to have > 1 point per row)
sfosmairports$geometry[1]
# Geometry set for 1 feature 
# geometry type:  POINT
# dimension:      XY
# bbox:           xmin: 29.73655 ymin: -2.595473 xmax: 29.73655 ymax: -2.595473
# geographic CRS: WGS 84
# POINT (29.73655 -2.595473)

sfairports$geometry[1]
# Geometry set for 1 feature 
# geometry type:  MULTIPOINT
# dimension:      XY
# bbox:           xmin: 30.13936 ymin: -1.970528 xmax: 30.13936 ymax: -1.970528
# geographic CRS: WGS 84
# MULTIPOINT ((30.13936 -1.970528))


# comparing the two files
mapview::mapview(list(sfosmairports, sfairports), col.regions=list('red','blue'))

# airport names
sfairports$NAME
#[1] "KIGALI"  "GISENYI" "KAMEMBE"
sfosmairports$name
#[1] "Butare Airport" "Nemba Airport"  "Gabiro Airport" NA               NA               NA      

# so the airports data from HDX reportedly from RNIS doesn't have Goma airport
# and the data from HDX from OSM doesn't have Gisenyi or Kigali
# 3 helipads don't have names

# an exercise could be to make a complete airports dataset for rwanda

# to get a dataset based on both
# we can 'bind' the rows together using rbind, but first they need to have the same columns
# if we try to bind them together when they have different columns an error will be generated

sfairportsboth <- rbind(sfosmairports, sfairports)
#Error: arguments have different crs

# aha this shows that they have different crs (Coordinate Reference System)
sf::st_crs(sfosmairports)
# Coordinate Reference System:
#   User input: WGS 84 etc.
sf::st_crs(sfairports)
# Coordinate Reference System:
#   User input: unnamed etc.

# here seems that they should have the same CRS (i.e. the most common one ESPSG 4326)
# can set the one that is not specified from the other
sf::st_crs(sfairports) <- sf::st_crs(sfosmairports)

#once the crs's match then get the error that the numbers of columns don't match
#Error in rbind.data.frame(...) : numbers of columns of arguments do not match

# dplyr::bind_rows does similar job allows binding and fills in empty columns with NAs
sfairportsboth <- dplyr::bind_rows(sfosmairports, sfairports)

# we can select & rename columns before the binding to make the final output cleaner & more useful
# we want to have the two files with the same column names with useful information in

# mutate allows us to rename columns, select to restrict columns

sfairports_tobind <- sfairports %>% 
                     mutate(name=NAME) %>%
                     mutate(type=FEATURE)  %>%
                     select(c(name,type,geometry))

sfosmairports_tobind <- sfosmairports %>% 
                     mutate(type=aeroway)  %>%
                     select(c(name,type,geometry))

#now bind to get a neat result with 9 rows and 3 columns
sfairportsboth <- dplyr::bind_rows(sfosmairports_tobind, sfairports_tobind)

str(sfairportsboth)
#Classes ‘sf’ and 'data.frame':	9 obs. of  3 variables:

# we can view an interactive map where the legend (using zcol) is set to the type, and labels are set to name
mapview::mapview(sfairportsboth, zcol='type', label='name')

# you could further edit values to make the data more consistent


```

```{r, line-data-airports, eval=TRUE, warning=FALSE}

# rwanda main roads
# https://data.humdata.org/dataset/rwanda-road-network
# click to download main roads (it may give a download warning), unzip all files

filename <- "C:\\Users\\andy.south\\Google Drive\\afrimapr-gdrive\\training\\rwanda-training\\rwanda-data\\rwa_trs_roads_osm.shp"

# read in a shapefile (shp) with sf package
sfmroads <- sf::st_read(filename)

str(sfmroads)
# 1080 rows, 21 columns including one for geometry

# Classes ‘sf’ and 'data.frame':	1080 obs. of  21 variables:
# ...
#  $ onme      : chr  "NR4" "NR4" "NR4" "NR4" ...
#  $ rtenme    : chr  "KN 5 Rd" NA "KN 5 Rd" NA ...
#  $ ntlclass  : chr  "trunk" "trunk" "trunk" "trunk" ...
#  $ fclass    : int  2 2 2 2 2 2 2 2 2 2 ...
#  $ numlanes  : int  2 NA 2 NA NA NA 2 NA 2 NA ...
# ...
#  $ geometry  :sfc_LINESTRING of length 1080; first list element:  'XY' num [1:16, 1:2] 30.1 30.1 30.1 30.1 30.1 

# have a quick look at a map
mapview::mapview(sfmroads)

# looking at the first row of the geometry column
# there are multiple pairs of coordinates because the file contains lines rather than points 

sfmroads$geometry[1]
# Geometry set for 1 feature 
# geometry type:  LINESTRING
# dimension:      XY
# bbox:           xmin: 30.12028 ymin: -1.962854 xmax: 30.12512 ymax: -1.961005
# geographic CRS: WGS 84
# LINESTRING (30.12028 -1.961005, 30.12065 -1.961...

# we can make the interactive map more informative as we did before using columns from the data 
# look at the data in RStudio to choose which columns contain most useful data 
# either click on the object name in the Environment tab top right or type :  
View(sfmroads)

# to add a legend using zcol, and hover labels using label
mapview::mapview(sfmroads, zcol='ntlclass', label='onme')




```