---
title: "Can open South Sudan Health facility location data help inform covid vaccine distribution ?"
#output: html_document
author: "Andy South @afrimapr"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: blue
always_allow_html: true
---

DRAFT

[Code for this document](https://github.com/afrimapr/afrimapr_dev/blob/master/compare_moh_southsudan.Rmd)    
[Packaged MoH Health facility data](https://github.com/afrimapr/afrimapr_dev/tree/master/southsudan/data-out) (with added Tier_name classification - see below)

# Summary

This draft report is part of a [pilot project run by MapAction](https://docs.google.com/document/d/1Rh9wCqHvJX8GLG68OAj3CzEMyvDXo-dc2rt7GTUQY6g/preview) funded through a private donor.

The purposes of the pilot project are:

* an immediately useful country data readiness package, documented and publicly available;
* to be in a better position to respond extremely quickly to COVAX’s needs; 
* and to build learning and community practice around the process to inform future work.

This report focuses on the availability of health facility data to inform covid vaccine distribution. It aims to inform the choice of data to include in the data package created by the pilot project.

The initial (draft) conclusion is that the data downloadable from the Ministry of Health website at [https://www.southsudanhealth.info/](https://www.southsudanhealth.info/) (**2889** recorded facilities) would be the most appropriate to include in the pilot data package. This is despite those data being dated and superceded locally by a cleaner dataset visible in the [South Sudan Health Service Functionality (HSF) Dashboard](https://southsudanhsf.shinyapps.io/hsf_dashboard/). The HSF data has fewer facilities (**1890**) with many more attributes, processed in collaboration between WHO and the Ministry of Health. Unfortunately, we do not have permission to use and/or redistribute these latter data, despite reaching out.   

The data stored in OpenStreetMap/healthsites.io are much sparser (**156**). We recommend improving the data in OpenStreetMap/healthsites.io as a longer term solution. For this project it is unlikely to be useful to combine the two datasets, partly because of technical difficulties in doing so, and partly because, when there are disagreements, there is no clear rationale for establishing which source is likely to be the more correct.   

If we were able to gain permission to access and use the data displayed in the [South Sudan Health Service Functionality Dashboard](https://southsudanhsf.shinyapps.io/hsf_dashboard/) (and summarised below) using them would be much better.     

\newpage

# Exploring available Health facility data

The following sources of health facility locations for South Sudan are available.

1. MoH   : downloaded from Ministry of Health website [https://www.southsudanhealth.info/](https://www.southsudanhealth.info/)
1. HSF   : [South Sudan Health Service Functionality Dashboard](https://southsudanhsf.shinyapps.io/hsf_dashboard/) (locations not downloadable, licensing uncertain). A collaboration between MoH & WHO.
1. OSM   : healthsites.io in OpenStreetMap
1. Kemri : collated data from published paper Maina(2019) Kemri, hosted by WHO GMP
1. HDX   : latest version on Humanitarian Data eXchange, labelled MoH 2016

For now I will focus mostly on the top three, because the latter two are older.

Table 1. Numbers of facility locations per source.

source  | num facilities | num hospitals | coords? | useful attributes
------- | ------ | ----- | ------------ | -----------
1. MoH     |  2889 | 130 | coords (2553 with, 336 without) |  only Facility codes e.g. FC05040202
3. HSF | 1890 | 82 | not downloadable, (1671 with coords, 219 without) | lots e.g. functionality, staff & bed numbers
2. OSM    | 432 | 72 | coords | (few complete) beds,staff_doctors,staff nurses
4. Kemri | 1734 | 40 | coords | no attributes
5. HDX | 1464 | 53 | coords | codes & functional status

The downloadable MoH data do not seem to have changed in the last year.

Both the MoH and HSF websites have contact emails indicating that they will answer queries. We reached out to both twice in April May 2021 and received no reply. We also emailed senior members of the WHO GIS team and received no reply.

The Functionality Dashboard says this about the data displayed there : "Facility data for this bulletin was kindly contributed by the Health Cluster partners, Health Pooled Fund (HPF), the International Committee of the Red Cross (ICRC), Management Sciences International (MSI), UNICEF, and UNHCR. Additional service availability data was extracted from the Ministry of Health DHIS2 platform for facilities without an implementing partner report. Most data was reported as of 31 March 2020, but the report date for each facility is indicated by hovering over a facility marker on the map. Implementing partners self-report on a quarterly basis service availability based predefined, standardized service availability data points and definitions. WHO merges the respective data sets and assigns a functionality level to each facility based on the criteria below."

Thus the HSF data seem to be the most comprehensive and up-to-date, but are not easily downloadable, do not have coordinates and do not have licencing information.

I did copy & paste the HSF data from the 19 separate pages so that I can summarise it in the figures below. I am not sure whether we are in a position to share these further without any permission from WHO or MoH.

I could try to join the data from HSF with that from the MoH website to add coordinates to the functionality data (and conversely to add attributes to the MoH data), but I'm not sure of the value of this given that it is re-doing something that someone in WHO/MoH has done already.

The MoH website also has associated attribute data (e.g. staff numbers), that can be accessed by clicking on a facility entry, but these data do not appear in the download, so this is not a feasible method for obtaining attributes for large numbers of facilities. In addition the licensing for these data is not stated. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(afrihealthsites)
library(knitr) #for kable
library(dplyr)
library(ggplot2)
library(patchwork) #to arrange plots
library(sf)
library(afriadmin)

```



```{r read-non-moh-data, eval=TRUE, include=FALSE, warning=FALSE}

# moh data is read in next chunk because coords need to be converted before reading into sf

filehsf <- "data-raw\\2021-05-ssd-hsfdashboard-copy.csv"
dfhsf <- read.csv(filehsf, as.is=TRUE, check.names=FALSE)
#Beware first column is not named, if you don't name it can get this error from dplyr later
#Error in initialize(...) : attempt to use zero-length variable name
names(dfhsf)[1] <- "ID" 

filehdx <- "data-raw\\list_hf_ss_moh_update_v1_20160404.csv"
dfhdx <- read.csv(filehdx, as.is=TRUE, check.names=FALSE)

sfosm <- afrihealthsites("south sudan", datasource="healthsites", plot=FALSE)


#as.data.frame(names(dfhsf))
# "OPD" "EPI"                      "ANC"                      "PNC"                      "SBA"                     
# [13] "BEmONC"                   "CEmONC"                   "FP"                       "PMTCT"                   
# [17] "ART"                      "TB"                       "BHI"                      "Inpatient - Paeds"       
# [21] "Inpatient - Adult"        "Basic surgery"            "MHPSS"                    "SGBV"                    
# [25] "NCD"                      "Nut. OTP"                 "Nut. TSFP"                "Nut. SC"                 
# [29] "Laboratory"               "Disability"               "# Clinical staff"         "Essential meds stockout" 
# [33] "FP stockout"              "Vaccine stockout"         "Ambulance access"         "# Inpatient beds - adult"
# [37] "# Inpatient beds - paeds" "Electricity"              "Latrine"                  "Dedicated mobile phone"  
# [41] "EPI refrigerator"         "Protected water source"

```

The HSF data has the richest attributes with these column names :

```{r, eval=TRUE, include=TRUE, echo=FALSE, warning=FALSE, comment=NA}

names(dfhsf)
#kable(names(dfhsf))

```


```{r, eval=TRUE, include=FALSE, warning=FALSE}

# read in files
filemoh <- "data-raw\\south_sudan_facility_info_2020-05-13.csv"
dfssd <- read.csv(filemoh, as.is=TRUE, check.names=FALSE)

#ssudan has a single location column containing coords divided by commas as strings 
# fair few coords at null island 0,0
# seemingly some NAs, or maybe problem with me trying to pass single Location rather than 2 columns
#which(is.na(dfssudan$Location))
#sfssudan <- sf::st_as_sf(dfssudan, coords='Location')

#dfs <- head(dfssudan)

#divide into 2 columns
dfcoords <- as.data.frame(stringr::str_split(dfssd$Location,", ",simplify=TRUE), stringsAsFactors = FALSE)
#change to numeric
dfcoords <- lapply(dfcoords,as.numeric)
#name columns
#names(dfcoords) <- c("Longitude", "Latitude")
names(dfcoords) <- c("Latitude", "Longitude")
#bind back onto df
dfssd <- cbind(dfssd, dfcoords)

sfmoh <- sf::st_as_sf(dfssd, coords=c("Longitude", "Latitude"), crs = 4326)

#mapview::mapview(sfmoh)

# I can add a 4 Tier classification here
# add Tier 0 for unknown
# TODO how best to do lookup ?
# I'm tired this is a daft way of doing but works
# I could put something into a function in afrihealhsites instead
unique(sfmoh$type) 
#  [1] "Other"                        "Primary Health Care Centre"   "Primary Health Care Unit"    
#  [4] "County Hospital"              "County Health Department"     "State Hospital"              
#  [7] "Specialized Hospital/Clinic"  "Teaching Hospital"            "Health Training Institutions"
# [10] "Ministry of Health"    
unique(dfhsf$`Facility type`) 

# PHCC, PHCU, Hospital, Mobile  

#should "Specialized Hospital/Clinic" go into Tier0 unknown or Tier4 ?
special <- dplyr::filter(sfmoh, type=="Specialized Hospital/Clinic")
#29 mobile clinics, camp clinics, eye clinics, IDP clinic, children hospital


# note that types not passed to 1 tier go into tier0 unknown
tiers <- afrihealthsites::tiers4(sfmoh$type,
                                 tier1=c("Primary Health Care Unit"),
                                 tier2=c("Primary Health Care Centre"),
                                 tier3=c("County Hospital"),
                                 tier4=c("State Hospital", "Teaching Hospital"),
                                 to_return = "Tier_name"
                                 )

sfmoh$Tier_name <- tiers

unique(sfmoh$Tier_name)
#Tier0 unknown, Tier2 health centre, Tier1 health post, Tier3 provincial hospital, Tier4 central hospital   

#save
save(sfmoh, file="sfmoh.rda")

#also write this as a geopackage
filename <- r"(data-out/ssd-moh-afrimapr-2021-05-28.gpkg)" #windows safe paths
sf::write_sf(sfmoh, filename)


```

```{r, eval=TRUE, include=FALSE, warning=FALSE}

#get admin boundaries because need to use them to subset OSM data
sfadm2county <- afriadmin("south sudan",level = 2,datasource = "geoboundaries",plot=FALSE)
sfadm1state <- afriadmin("south sudan",level = 1,datasource = "geoboundaries",plot=FALSE)

sfadm2juba <- dplyr::filter(sfadm2county, shapeName=="Juba")


```


```{r map_moh_ssudan, eval=FALSE, include=FALSE, warning=FALSE}

# 2889 obs

# just view MoH data
sfmoh <- afrihealthsites("South Sudan", datasource=ssudan_file, lonlat_columns = "Location")

sfmoh <- afrihealthsites("South Sudan", datasource=sfmoh,                              
                         type_column = 'type',
                         label_column = 'Facility')


# plot moh vs kemri on a map
# afrihealthsites::compare_hs_sources("south sudan", datasources=c('who',ssudan_file))

# most overlap but not total


```


```{r moh_type_frequencies, echo=FALSE, warnings=FALSE, asis=TRUE, fig.width=9, fig.height=6}

# Plot counts of facility types from MoH data

# pass the sf object directly
ggmoh <- facility_types("south sudan", 
                        datasource=sfmoh,
                        type_column = "type",
                        plot_title=paste("Facility types, MoH www.southsudanhealth.info. Total:", nrow(sfmoh) ),
                        plot=FALSE)

#save plot for patchwork in later chunk to align figs
#plot(ggssd)



```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=4}

# facility data (no coords) copied from WHO/MOH health service functionality (hsf) dashboard


# read in files, now done earlier
#filehsf <- "data-raw\\2021-05-ssd-hsfdashboard-copy.csv"
#dfhsf <- read.csv(filehsf, as.is=TRUE, check.names=FALSE)
#Beware first column is not named, if you don't name it can get this error from dplyr later
#Error in initialize(...) : attempt to use zero-length variable name
#names(dfhsf)[1] <- "ID" 

gghsf <- facility_types("south sudan", 
                        datasource=dfhsf,
                        type_column = "Facility type",
                        lonlat_columns = NULL,
                        plot_title=paste("Facility types, Health Service Functionality dashboard (no coords). Total:", nrow(dfhsf)),
                        plot=FALSE)

#save plot for patchwork in later chunk to align figs
#plot(gghsf)

```

```{r osm_type_frequencies, echo=FALSE, warnings=FALSE, asis=TRUE, fig.width=9, fig.height=6}

# Plot counts of facility types from OSM

ggosm <- facility_types("south sudan", 
                        datasource="healthsites",
                        type_filter=c('clinic','doctors','hospital'),
                        #plot_title=paste("Facility types, MoH www.southsudanhealth.info. Total:", nrow(sfmoh) ),
                        plot=FALSE)

#save plot for patchwork in later chunk to align figs
#plot(ggosm)



```


```{r, eval=TRUE, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=14}

# facility data from HDX

gghdx <- facility_types("south sudan", 
                        datasource=dfhdx,
                        type_column = "HF_TYPE",
                        lonlat_columns = NULL,
                        plot_title=paste("Facility types, latest HDX (from MoH 2016). Total:", nrow(dfhdx)),
                        plot=FALSE)
#plot(gghdx)

#set xmax to be the same for all plots
gg1 <- ggmoh
gg2 <- gghsf
gg3 <- gghdx
gg4 <- ggosm

#hack to find max xlim for each object
#TODO make this less hacky ! it will probably fail when ggplot changes
max_x1 <- max(ggplot_build(gg1)$layout$panel_params[[1]]$x$continuous_range)
max_x2 <- max(ggplot_build(gg2)$layout$panel_params[[1]]$x$continuous_range)
max_x3 <- max(ggplot_build(gg3)$layout$panel_params[[1]]$x$continuous_range)
max_x4 <- max(ggplot_build(gg4)$layout$panel_params[[1]]$x$continuous_range)
#set xmax for all plots to this
max_x <- max(max_x1,max_x2,max_x3,max_x4, na.rm=TRUE)
gg1 <- gg1 + xlim(c(0,max_x))
gg2 <- gg2 + xlim(c(0,max_x))
gg3 <- gg3 + xlim(c(0,max_x))
gg4 <- gg4 + xlim(c(0,max_x))

#align 3 figs using patchwork, set height to num types so that bar widths are the same 
#ggmoh / gghsf / gghdx /ggosm + plot_layout(heights = c(length(unique(sfmoh$type)),
gg1 / gg2 / gg3 / gg4 + plot_layout(heights = c(length(unique(sfmoh$type)),                 
                                                       length(unique(dfhsf$`Facility type`)),
                                                       length(unique(dfhdx$HF_TYPE)),
                                                       3
                                                       ))

```

```{r moh_missing_coords, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=5}


sfmohnocoords <- sfmoh[which(sfmoh$Location=="0.00000, 0.00000"),] 
sfmohcoords <- sfmoh[which(sfmoh$Location!="0.00000, 0.00000"),] 

# types for those without coords
ggssd <- facility_types("south sudan", 
                        datasource=sfmohnocoords,
                        type_column = "type",
                        plot_title="336 Facilities in the MoH data with no coordinates",
                        plot=FALSE)
plot(ggssd)


```

I've added a 5 Tier classification (Tier_name) that any local classification can be converted to. 

* Tier0 unknown 
* Tier2 health centre 
* Tier1 health post 
* Tier3 provincial hospital 
* Tier4 central hospital. 

This similar to the 4 tier classification used in the [Falchetta(2020)](https://www.pnas.org/content/117/50/31760) paper, it just adds a Tier0 for unknown. 

```{r moh_4tiers, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=4}


# types for those without coords
ggssd_tiers <- facility_types("south sudan", 
                        datasource=sfmoh,
                        type_column = "Tier_name",
                        plot_title="MoH data reclassified to 4 Tiers of Falchetta(2020)",
                        plot=FALSE)
plot(ggssd_tiers)


```







```{r, eval=TRUE, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=14}

# hospitals alone

#this selects just hospitals from MoH (130) & HSF (82) & OSM () data
sfmoh_hosp <- dplyr::filter(sfmoh, grepl("Hospital", type))
dfhsf_hosp <- dplyr::filter(dfhsf, grepl("Hospital", `Facility type`))
sfosm_hosp <- dplyr::filter(sfosm, grepl("hospital", amenity))

```

\newpage

# Subsetting Hospitals in Juba county for initial detailed comparison

This shows that, although some are shared, hospital names differ between the 3 sources.

It gives an idea of the magnitude of the problem of trying to combine data sources, and points to the potential folly of trying to repeat data cleaning that has already been done to arrive at the Health Service Functionality (hsf) dataset.

Even when names are shared between data sources they may differ slightly e.g. 	
moh : El Sabbah Children Hosp    
hsf : El-Sabah Children Hospital

The OpenStreetMap data show examples of : 

1. un-named facilities (row1)
1. facilities that appear not to be hospitals (e.g. row7 primary health care unit, row8 high court & row14 small clinic)
1. facilities now closed (row18)

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, asis=TRUE, fig.width=9, fig.height=14}


#Tricky because MoH data uses Jubek state. Jubek State was a state in South Sudan that existed between 2 October 2015 and 22 February 2020. It was based on Juba county.


#7 hospitals in MoH
sfmoh2 <- dplyr::filter(sfmoh, grepl("Hospital", type) & County == "Juba County")
#13 Hospitals in HSF
dfhsf2 <- dplyr::filter(dfhsf, grepl("Hospital", `Facility type`) & County == "Juba")

# OSM need to do spatial operation because no column for county
# aha can use afrihealthsites function
# initially it only worked with admin1 due to bug in afrihealthsites
#19 hospitals in OSM
sfosm2 <- afrihealthsites("south sudan", datasource="healthsites", plot=FALSE,
                          hs_amenity=c('hospital'),
                          admin_level=2, admin_names=c('Juba') )
                          #admin_level=1, admin_names=c('Central Equatoria') )

# this suggests there should be OSM hospitals within Juba county
#mapview::mapview(list(sfosm_hosp, sfadm2county))

x <- sort(sfmoh2$Facility)
y <- sort(dfhsf2$Facility)
z <- sort(sfosm2$name)
n <- max(length(x), length(y), length(z))
length(x) <- length(y) <- length(z) <- n

dfnames <- cbind(row=1:n,
                 moh=x,
                 hsf=y,
                 osm=z)

options(knitr.kable.NA = '')

kable(dfnames, caption="Juba county hospitals from different data sources")

# dfnames <- data.frame(moh=sort(sfmoh2$Facility),
#                       hsf=sort(dfhsf2$Facility),
#                       osm=sort(sfosm2$Facility))
# dfnames <- cbind(moh=sort(sfmoh2$Facility),
#                       hsf=sort(dfhsf2$Facility),
#                       osm=sort(sfosm2$name))



                      
# kable(sfmoh2$Facility)
# kable(dfhsf2$Facility)
# kable(sfosm2$Facility)

#align 3 figs using patchwork, set height to num types so that bar widths are the same 
# ggmoh / gghsf / gghdx /ggosm + plot_layout(heights = c(length(unique(sfmoh$type)),
#                                                        length(unique(dfhsf$`Facility type`)),
#                                                        length(unique(dfhdx$HF_TYPE)),
#                                                        3
#                                                        ))

```

\newpage

# Exploring potentially useful attributes from Health Service Functionality data

(Note, it is not clear whether we have permission to use these data in this way).

```{r hsf_functionality, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=5}

dfhsf$Functionality[ which(dfhsf$Functionality == "") ] <- "unknown"

dfhsf$Functionality <- factor(dfhsf$Functionality, levels = c(	
"Highly Functional", "Moderate", "Limited", "unknown", "Non-Functional"))

gg <- ggplot2::ggplot(dfhsf, aes(y = `Facility type`, fill = `Functionality`))

gg <- gg + geom_bar(show.legend=TRUE) +
      labs(title="HSF data, Functionality by Facility type") +
      scale_fill_manual(values=c("green4","green3","darkorange", "white","darkred")) +
      theme_minimal() 
      
gg


```

```{r hsf_electricity, echo=FALSE, warning=FALSE, asis=TRUE, fig.width=9, fig.height=5}

dfhsf$Electricity[ which(dfhsf$Electricity == "") ] <- "unknown"

dfhsf$Electricity <- factor(dfhsf$Electricity, levels = c("Available", "unknown", "Not Available"))

gg <- ggplot2::ggplot(dfhsf, aes(y = `Facility type`, fill = `Electricity`))

gg <- gg + geom_bar(show.legend=TRUE) +
      labs(title="HSF data, Electricity availability by Facility type") +
      scale_fill_manual(values=c("darkgreen", "darkorange", "darkred")) +
      theme_minimal() 
      
gg


```